Here I am cleaning up my results code because it is getting messy

Packages
```{r}
library(dplyr)
library(tidyverse)
library(strucchange)
library(readr)
library(rootSolve)
library(readxl)
library(MGnifyR)
library(biomformat)
library(island)
library(readxl)
library(DescTools)
library(stringr)
library(vegan)
library(data.table)
library(adespatial)
library(codyn)
library(RobustLinearReg)
```


Data
```{r}
data_.0001 <- read_rds("Studies_scripts//final_flat_.0001.rds")

num_timepoints_.0001 <- lapply(data_.0001, function(study){
  return(length(colnames(study))-9)
})
```


Cleaning Data
1. removing contaminants
2. replacing all numbers in the "family" column with blanks because these are not resolved to family


SHOULD I REPLACE ALL BLANK FAMILIES WITH UNKNOWNS SO I CAN INCLUDE ALL DATA?
NO!!!! Think about it - the ones that aren't resolved to family are grouping multiple families together so it groups these levels together. (think about why you are looking at family).
- justification for family as level: level that includes as much diversity and as many samples across the board as possible (some don't resolve all the way to genus, order was too broad not enough diversity)
- potentially do genus as well????????

3. Creating a data frame with no rows that will produce artifacts with the model (rows with fewer than 2 transitions), while retaining information about those rows in a separate data table.
```{r}
##Trying to remove chloroplasts, mitochondria, and non-microbial DNA
remove_contaminants <- function(df, tax_cols = 1:9, patterns = c("Chloroplast", "Mitochondria", "Metazoa", "Plantae")) {
  # Create regex pattern
  pattern <- paste(patterns, collapse = "|")
  
  # Subset taxonomy columns
  tax_data <- df[, tax_cols]
  
  # Create a logical vector for rows to keep
  # TRUE if none of the taxonomy columns contain the pattern
  keep <- !apply(tax_data, 1, function(row) {
    any(grepl(pattern, row, ignore.case = TRUE))
  })
  
  # Return the cleaned dataframe
  df[keep, ]
}

data_.0001_clean <- lapply(data_.0001, remove_contaminants)
#removing numbers from the family column
data_.0001_clean <- lapply(data_.0001_clean, function(df) {
  # Replace if 'family' is only digits (as string or number)
  df$family <- as.character(df$family)  # ensure it's character
  df$family[grepl("\\d", df$family)] <- ""
  return(df)
})

#removing artifacts
analyze_transitions <- function(vec) {
  pa <- as.numeric(vec)
  transitions <- sum(diff(pa) != 0, na.rm = TRUE)

  if (all(pa == 1)) return("all_1s")
  if (all(pa == 0)) return("all_0s")
  if (transitions == 1) {
    first_change <- which(diff(pa) != 0)[1]
    from <- pa[first_change]
    to <- pa[first_change + 1]
    if (from == 0 && to == 1) return("0_to_1")
    if (from == 1 && to == 0) return("1_to_0")
  }

  return(NA)
}

process_otu_list <- function(otu_list) {
  # Output list of filtered tables
  filtered_list <- list()
  # Data frame of removed rows and reasons
  removal_log <- data.frame()

  for (i in seq_along(otu_list)) {
    df <- otu_list[[i]]
    tax_cols <- df[, 1:9]
    pa_cols <- df[, 10:ncol(df)]

    # Analyze each row
    removal_reasons <- apply(pa_cols, 1, analyze_transitions)

    # Keep rows without a removal reason
    keep_idx <- is.na(removal_reasons)
    filtered_list[[i]] <- df[keep_idx, ]

    # Store the removed rows and reasons
    if (any(!keep_idx)) {
      removed <- df[!keep_idx, 1:9]
      removed$reason <- removal_reasons[!keep_idx]
      removed$source_table <- names(otu_list)[i]
      removal_log <- rbind(removal_log, removed)
    }
  }

  return(list(filtered_otus = filtered_list, removed_info = removal_log))
}

removed_artifacts <- process_otu_list(data_.0001_clean)
tables_no_artifacts <- removed_artifacts$filtered_otus
names(tables_no_artifacts) <- names(data_.0001_clean)
removed_no_artifacts <- removed_artifacts$removed_info

```

Running the colonization-extinction model on these datasets.
```{r}
get_ce_params_fam <- function(otu_table) {
  ran <- 10:length(colnames(otu_table))
  fam <- island::irregular_single_dataset(otu_table, ran, c = 0.00001, e = 0.00001,
                                          column = "family", n = 1)
  
  fam_c <- fam$c[-1]
  fam_e <- fam$e[-1]
  
  print("done")
  
  return(list("fam_c" = fam_c, "fam_e" = fam_e, "details" = fam))
}

suppressWarnings({
  fam_.0001 <- lapply(data_.0001_clean, get_ce_params_fam)
})

suppressWarnings({
  fam_no_art <- lapply(tables_no_artifacts, get_ce_params_fam)
})
```


Getting Rho, variance, and number of unique families for each dataset
```{r}
max_results.0001 <- lapply(fam_.0001, function(fam_result){
  fam_result$famrichness <- length(fam_result$fam_c)
  fam_result$total_richness <- sum(fam_result$details$N)
  fam_result$distinct_families <- length(unique(fam_result$fam_c))
  return(fam_result)
})

max_results_no_art <- lapply(fam_no_art, function(fam_result){
  fam_result$famrichness <- length(fam_result$fam_c)
  fam_result$total_richness <- sum(fam_result$details$N)
  fam_result$distinct_families <- length(unique(fam_result$fam_c))
  return(fam_result)
})

compute_cp_rho <- function(ce_data) {
  if(length(ce_data$fam_c) > 1){
    result <- cor.test(ce_data$fam_c, 1/ce_data$fam_e, method = "spearman")
    n <- length(ce_data$fam_c)
    result$n <- n
    result$total_richness <- ce_data$total_richness
    result$distinct_families <- ce_data$distinct_families
  }
  else{
    result <- list()
    result$estimate <- NA
    result$p.value <- NA
    n <- length(ce_data$fam_c)
    result$n <- n
    result$total_richness <- ce_data$total_richness
    result$distinct_families <- ce_data$distinct_families
  }
  return(result)
}

cp_rho_list.0001 <- lapply(max_results.0001, compute_cp_rho)
cp_rho_list_no_art <- lapply(max_results_no_art, compute_cp_rho)

result.0001_df <- do.call(rbind, lapply(names(cp_rho_list.0001), function(id) {
  item <- cp_rho_list.0001[[id]]
  data.frame(
    ID = id,
    rho = item$estimate,
    p_value = item$p.value,
    n = item$n,
    richness = item$total_richness,
    distinct_families = item$distinct_families,
    stringsAsFactors = FALSE
  )
}))

result.0001_df$num_timepoint <- num_timepoints_.0001

result_no_art_df <- do.call(rbind, lapply(names(cp_rho_list_no_art), function(id) {
  item <- cp_rho_list_no_art[[id]]
  data.frame(
    ID = id,
    rho = item$estimate,
    p_value = item$p.value,
    n = item$n,
    richness = item$total_richness,
    distinct_families = item$distinct_families,
    stringsAsFactors = FALSE
  )
}))

result_no_art_df$num_timepoint <- num_timepoints_.0001

```

Cleaning up the data
```{r}
#removing extra data (duplicate samples) and nonhuman samples
view(result.0001_df)
biomes_to_keep <- result.0001_df[c(1,2,38,42,699:733),]
view(biomes_to_keep)
final_biome_list.0001 <- result.0001_df %>%
  filter(!str_detect(ID, "extra|Cat|House|Dog|Bgtw"))
##Adding back corrected mistakes
final_biome_list.0001 <- rbind(final_biome_list.0001, biomes_to_keep)
final_biome_list.0001 <- final_biome_list.0001 %>%
  filter(!str_detect(ID, "MT5147_Bgtw3.M_gut.extra"))

#####total biome/subject combinations!!!!
final_biome_list.0001$rho <- as.character(final_biome_list.0001$rho)
final_biome_list.0001$rho[which(str_detect(final_biome_list.0001$rho, "-1")==TRUE)] <- NA
final_biome_list.0001$rho <- as.numeric(final_biome_list.0001$rho)
final_biome_list.0001$rho[which(final_biome_list.0001$rho==-1)] <- NA
final_biome_list.0001$rho[which(final_biome_list.0001$rho==1)] <- NA
final_biome_list.0001$z <- lapply(final_biome_list.0001$rho, FisherZ)
final_biome_list.0001$v <- 1/(final_biome_list.0001$n-3)
final_biome_list.0001 <- as.data.table(final_biome_list.0001)
final_biome_list.0001 <- final_biome_list.0001[, c("study_ID", "subject_id", "biome_ID") := tstrsplit(ID, "_")]
final_biome_list.0001 <- as.data.frame(final_biome_list.0001)
final_biome_list.0001$z <- as.numeric(final_biome_list.0001$z) 
final_biome_list.0001$num_timepoint <- as.numeric(final_biome_list.0001$num_timepoint)
final_biome_list.0001 <- final_biome_list.0001[-c(1:2),]

write.csv(final_biome_list.0001, "final_biome_list.0001.csv")
view(final_biome_list.0001)

##############################################################################

###Doing the same thing for no_art
#removing extra data (duplicate samples) and nonhuman samples
view(result_no_art_df)
biomes_to_keep_no_art <- result_no_art_df[c(1,2,38,42,699:733),]
view(biomes_to_keep)
final_biome_list_no_art <- result_no_art_df %>%
  filter(!str_detect(ID, "extra|Cat|House|Dog|Bgtw"))
##Adding back corrected mistakes
final_biome_list_no_art <- rbind(final_biome_list_no_art, biomes_to_keep_no_art)
final_biome_list_no_art <- final_biome_list_no_art %>%
  filter(!str_detect(ID, "MT5147_Bgtw3.M_gut.extra"))

#####total biome/subject combinations!!!!
final_biome_list_no_art$rho <- as.character(final_biome_list_no_art$rho)
final_biome_list_no_art$rho[which(str_detect(final_biome_list_no_art$rho, "-1")==TRUE)] <- NA
final_biome_list_no_art$rho <- as.numeric(final_biome_list_no_art$rho)
final_biome_list_no_art$rho[which(final_biome_list_no_art$rho==-1)] <- NA
final_biome_list_no_art$rho[which(final_biome_list_no_art$rho==1)] <- NA
final_biome_list_no_art$z <- lapply(final_biome_list_no_art$rho, FisherZ)
final_biome_list_no_art$v <- 1/(final_biome_list_no_art$n-3)
final_biome_list_no_art <- as.data.table(final_biome_list_no_art)
final_biome_list_no_art <- final_biome_list_no_art[, c("study_ID", "subject_id", "biome_ID") := tstrsplit(ID, "_")]
final_biome_list_no_art <- as.data.frame(final_biome_list_no_art)
final_biome_list_no_art$z <- as.numeric(final_biome_list_no_art$z) 
final_biome_list_no_art$num_timepoint <- as.numeric(final_biome_list_no_art$num_timepoint)
final_biome_list_no_art <- final_biome_list_no_art[-c(1:2),]

view(final_biome_list_no_art)

write.csv(final_biome_list_no_art, "final_biome_list_no_art.csv")

```


Another thing I need to calculate for the two datasets separately is slopes
```{r}
compute_log_cp_relationship <- function(ce_data, study_id) {
  result <- tryCatch({
    if (length(ce_data$fam_c) > 1) {
      model <- lm(log(ce_data$fam_c) ~ log(1 / ce_data$fam_e))
      s <- summary(model)
      attr(s, "study_id") <- study_id
      s
    } else {
      NA
    }
  }, warning = function(w) {
    message(paste("Warning for:", study_id, "-", conditionMessage(w)))
    return(NA)
  }, error = function(e) {
    message(paste("Error for:", study_id, "-", conditionMessage(e)))
    return(NA)
  })
  return(result)
}

compute_log_pc_relationship <- function(ce_data, study_id) {
  result <- tryCatch({
    if (length(ce_data$fam_c) > 1) {
      c <- log(ce_data$fam_c)
      p <- log(1/ce_data$fam_e)
      model <- lm(p~c)
      s <- summary(model)
      attr(s, "study_id") <- study_id
      s
    } else {
      NA
    }
  }, warning = function(w) {
    message(paste("Warning for:", study_id, "-", conditionMessage(w)))
    return(NA)
  }, error = function(e) {
    message(paste("Error for:", study_id, "-", conditionMessage(e)))
    return(NA)
  })
  return(result)
}

cp_log_slopes.0001 <- mapply(
  compute_log_cp_relationship,
  ce_data = max_results.0001,
  study_id = names(max_results.0001),
  SIMPLIFY = FALSE
)
cp_log_slopes.0001$M0278_Infant1_gut$coefficients[2,1]

se_and_slope <- do.call(rbind,(lapply(cp_log_slopes.0001, function(result){
  if(!is.na(result[1])){
    slope <- result$coefficients[2,1]
    ste <- result$coefficients[2,2]
    slope_p <- result$coefficients[2,4]
  }else{
    slope <- NA
    ste <- NA
    slope_p <- NA
  }
  row = data.frame("slope" = slope, "ste" = ste, "slope_p" = slope_p)
  return(row)
})))
se_and_slope$ID <- rownames(se_and_slope)

merged.0001 <- se_and_slope %>%
  inner_join(x = se_and_slope, y = final_biome_list.0001, by = join_by(ID==ID))

pc_log_slopes.0001 <- mapply(
  compute_log_pc_relationship,
  ce_data = max_results.0001,
  study_id = names(max_results.0001),
  SIMPLIFY = FALSE
)

se_and_slope_reverse <- do.call(rbind,(lapply(pc_log_slopes.0001, function(result){
  if(!is.na(result[1])){
    slope <- result$coefficients[2,1]
    ste <- result$coefficients[2,2]
    slope_p <- result$coefficients[2,4]
  }else{
    slope <- NA
    ste <- NA
    slope_p <- NA
  }
  row = data.frame("pc_slope" = slope, "pc_ste" = ste, "pc_slope_p" = slope_p)
  return(row)
})))
se_and_slope_reverse$ID <- rownames(se_and_slope_reverse)

merged.0001 <- se_and_slope_reverse %>%
  inner_join(x = se_and_slope_reverse, y = merged.0001, by = join_by(ID==ID))

#########################################################
#for no_art
cp_log_slopes_no_art <- mapply(
  compute_log_cp_relationship,
  ce_data = max_results_no_art,
  study_id = names(max_results_no_art),
  SIMPLIFY = FALSE
)

se_and_slope_no_art <- do.call(rbind,(lapply(cp_log_slopes_no_art, function(result){
  if(!is.na(result[1])){
    slope <- result$coefficients[2,1]
    ste <- result$coefficients[2,2]
    slope_p <- result$coefficients[2,4]
  }else{
    slope <- NA
    ste <- NA
    slope_p <- NA
  }
  row = data.frame("slope" = slope, "ste" = ste, "slope_p" = slope_p)
  return(row)
})))
se_and_slope_no_art$ID <- rownames(se_and_slope_no_art)

merged_no_art <- se_and_slope_no_art %>%
  inner_join(x = se_and_slope_no_art, y = final_biome_list_no_art, by = join_by(ID==ID))

pc_log_slopes_no_art <- mapply(
  compute_log_pc_relationship,
  ce_data = max_results_no_art,
  study_id = names(max_results.0001),
  SIMPLIFY = FALSE
)

se_and_slope_reverse_no_art <- do.call(rbind,(lapply(pc_log_slopes_no_art, function(result){
  if(!is.na(result[1])){
    slope <- result$coefficients[2,1]
    ste <- result$coefficients[2,2]
    slope_p <- result$coefficients[2,4]
  }else{
    slope <- NA
    ste <- NA
    slope_p <- NA
  }
  row = data.frame("pc_slope" = slope, "pc_ste" = ste, "pc_slope_p" = slope_p)
  return(row)
})))
se_and_slope_reverse_no_art$ID <- rownames(se_and_slope_reverse_no_art)

merged_no_art <- se_and_slope_reverse_no_art %>%
  inner_join(x = se_and_slope_reverse_no_art, y = merged_no_art, by = join_by(ID==ID))

write.csv(merged.0001, "merged_data_nobeta.0001.csv")
write.csv(merged_no_art, "merged_data_nobeta_no_art.csv")

```

Now I am going to get all of this info on the genus level
```{r}
data_.0001_genus <- lapply(data_.0001_clean, function(df) {
  # Replace if 'family' is only digits (as string or number)
  df$family <- as.character(df$genus)  # ensure it's character
  df$family[grepl("\\d", df$genus)] <- ""
  return(df)
})

data_no_art_genus <- lapply(tables_no_artifacts, function(df) {
  # Replace if 'family' is only digits (as string or number)
  df$family <- as.character(df$genus)  # ensure it's character
  df$family[grepl("\\d", df$genus)] <- ""
  return(df)
})

get_ce_params_gen <- function(otu_table) {
  ran <- 10:length(colnames(otu_table))
  fam <- island::irregular_single_dataset(otu_table, ran, c = 0.00001, e = 0.00001,
                                          column = "genus", n = 1)
  
  fam_c <- fam$c[-1]
  fam_e <- fam$e[-1]
  
  print("done")
  
  return(list("gen_c" = fam_c, "gen_e" = fam_e, "details" = fam))
}

suppressWarnings({
  gen_.0001 <- lapply(data_.0001_genus, get_ce_params_gen)
})

suppressWarnings({
  gen_no_art <- lapply(data_no_art_genus, get_ce_params_gen)
})

max_results_gen.0001 <- lapply(gen_.0001, function(fam_result){
  fam_result$genrichness <- length(fam_result$gen_c)
  fam_result$total_richness <- sum(fam_result$details$N)
  fam_result$distinct_genera <- length(unique(fam_result$gen_c))
  return(fam_result)
})

max_results_gen_no_art <- lapply(gen_no_art, function(fam_result){
  fam_result$genrichness <- length(fam_result$gen_c)
  fam_result$total_richness <- sum(fam_result$details$N)
  fam_result$distinct_genera <- length(unique(fam_result$gen_c))
  return(fam_result)
})

compute_cp_rho_gen <- function(ce_data) {
  if(length(ce_data$gen_c) > 1){
    result <- cor.test(ce_data$gen_c, 1/ce_data$gen_e, method = "spearman")
    n <- length(ce_data$gen_c)
    result$n <- n
    result$total_richness <- ce_data$total_richness
    result$distinct_genera <- ce_data$distinct_genera
  }
  else{
    result <- list()
    result$estimate <- NA
    result$p.value <- NA
    n <- length(ce_data$gen_c)
    result$n <- n
    result$total_richness <- ce_data$total_richness
    result$distinct_genera <- ce_data$distinct_genera
  }
  return(result)
}

cp_rho_list_gen.0001 <- lapply(max_results_gen.0001, compute_cp_rho_gen)
cp_rho_list_gen_no_art <- lapply(max_results_gen_no_art, compute_cp_rho_gen)

result_gen.0001_df <- do.call(rbind, lapply(names(cp_rho_list_gen.0001), function(id) {
  item <- cp_rho_list_gen.0001[[id]]
  data.frame(
    ID = id,
    rho = item$estimate,
    p_value = item$p.value,
    n = item$n,
    richness = item$total_richness,
    distinct_genera = item$distinct_genera,
    stringsAsFactors = FALSE
  )
}))

result_gen.0001_df$num_timepoint <- num_timepoints_.0001

result_no_art_gen_df <- do.call(rbind, lapply(names(cp_rho_list_gen_no_art), function(id) {
  item <- cp_rho_list_gen_no_art[[id]]
  data.frame(
    ID = id,
    rho = item$estimate,
    p_value = item$p.value,
    n = item$n,
    richness = item$total_richness,
    distinct_genera = item$distinct_genera,
    stringsAsFactors = FALSE
  )
}))

result_no_art_gen_df$num_timepoint <- num_timepoints_.0001

#########################SLOPES########################################
compute_log_cp_relationship_gen <- function(ce_data, study_id) {
  result <- tryCatch({
    if (length(ce_data$gen_c) > 1) {
      model <- lm(log(ce_data$gen_c) ~ log(1 / ce_data$gen_e))
      s <- summary(model)
      attr(s, "study_id") <- study_id
      s
    } else {
      NA
    }
  }, warning = function(w) {
    message(paste("Warning for:", study_id, "-", conditionMessage(w)))
    return(NA)
  }, error = function(e) {
    message(paste("Error for:", study_id, "-", conditionMessage(e)))
    return(NA)
  })
  return(result)
}

compute_log_pc_relationship_gen <- function(ce_data, study_id) {
  result <- tryCatch({
    if (length(ce_data$gen_c) > 1) {
      c <- log(ce_data$gen_c)
      p <- log(1/ce_data$gen_e)
      model <- lm(p~c)
      s <- summary(model)
      attr(s, "study_id") <- study_id
      s
    } else {
      NA
    }
  }, warning = function(w) {
    message(paste("Warning for:", study_id, "-", conditionMessage(w)))
    return(NA)
  }, error = function(e) {
    message(paste("Error for:", study_id, "-", conditionMessage(e)))
    return(NA)
  })
  return(result)
}

cp_log_slopes_gen.0001 <- mapply(
  compute_log_cp_relationship_gen,
  ce_data = max_results_gen.0001,
  study_id = names(max_results_gen.0001),
  SIMPLIFY = FALSE
)

se_and_slope_gen.0001 <- do.call(rbind,(lapply(cp_log_slopes_gen.0001, function(result){
  if(!is.na(result[1])){
    slope <- result$coefficients[2,1]
    ste <- result$coefficients[2,2]
    slope_p <- result$coefficients[2,4]
  }else{
    slope <- NA
    ste <- NA
    slope_p <- NA
  }
  row = data.frame("slope" = slope, "ste" = ste, "slope_p" = slope_p)
  return(row)
})))
se_and_slope_gen.0001$ID <- rownames(se_and_slope_gen.0001)

merged_gen.0001 <- se_and_slope_gen.0001 %>%
  inner_join(x = se_and_slope_gen.0001, y = result_gen.0001_df, by = join_by(ID==ID))

pc_log_slopes_gen.0001 <- mapply(
  compute_log_pc_relationship_gen,
  ce_data = max_results_gen.0001,
  study_id = names(max_results_gen.0001),
  SIMPLIFY = FALSE
)

se_and_slope_reverse_gen.0001 <- do.call(rbind,(lapply(pc_log_slopes_gen.0001, function(result){
  if(!is.na(result[1])){
    slope <- result$coefficients[2,1]
    ste <- result$coefficients[2,2]
    slope_p <- result$coefficients[2,4]
  }else{
    slope <- NA
    ste <- NA
    slope_p <- NA
  }
  row = data.frame("pc_slope" = slope, "pc_ste" = ste, "pc_slope_p" = slope_p)
  return(row)
})))
se_and_slope_reverse_gen.0001$ID <- rownames(se_and_slope_reverse_gen.0001)

merged_gen.0001 <- se_and_slope_reverse_gen.0001 %>%
  inner_join(x = se_and_slope_reverse_gen.0001, y = merged_gen.0001, by = join_by(ID==ID))
#####filter by list of studies included
final_biome_list.0001$ID
merged_gen.0001 <- merged_gen.0001 %>% filter(ID %in% final_biome_list.0001$ID)

##################################################################
#slope for no artifacts
cp_log_slopes_gen_no_art <- mapply(
  compute_log_cp_relationship_gen,
  ce_data = max_results_gen_no_art,
  study_id = names(max_results_gen_no_art),
  SIMPLIFY = FALSE
)

se_and_slope_gen_no_art <- do.call(rbind,(lapply(cp_log_slopes_gen_no_art, function(result){
  if(!is.na(result[1])){
    slope <- result$coefficients[2,1]
    ste <- result$coefficients[2,2]
    slope_p <- result$coefficients[2,4]
  }else{
    slope <- NA
    ste <- NA
    slope_p <- NA
  }
  row = data.frame("slope" = slope, "ste" = ste, "slope_p" = slope_p)
  return(row)
})))
se_and_slope_gen_no_art$ID <- rownames(se_and_slope_gen_no_art)

merged_gen_no_art <- se_and_slope_gen_no_art %>%
  inner_join(x = se_and_slope_gen_no_art, y = result_no_art_gen_df, by = join_by(ID==ID))

pc_log_slopes_gen_no_art <- mapply(
  compute_log_pc_relationship_gen,
  ce_data = max_results_gen_no_art,
  study_id = names(max_results_gen_no_art),
  SIMPLIFY = FALSE
)

se_and_slope_reverse_gen_no_art <- do.call(rbind,(lapply(pc_log_slopes_gen_no_art, function(result){
  if(!is.na(result[1])){
    slope <- result$coefficients[2,1]
    ste <- result$coefficients[2,2]
    slope_p <- result$coefficients[2,4]
  }else{
    slope <- NA
    ste <- NA
    slope_p <- NA
  }
  row = data.frame("pc_slope" = slope, "pc_ste" = ste, "pc_slope_p" = slope_p)
  return(row)
})))
se_and_slope_reverse_gen_no_art$ID <- rownames(se_and_slope_reverse_gen_no_art)

merged_gen_no_art <- se_and_slope_reverse_gen_no_art %>%
  inner_join(x = se_and_slope_reverse_gen_no_art, y = merged_gen_no_art, by = join_by(ID==ID))
#####filter by list of studies included
final_biome_list_no_art$ID
merged_gen_no_art <- merged_gen_no_art %>% filter(ID %in% final_biome_list_no_art$ID)



##############cleanup#####################
cleanup <- function(biome_list) {
  biome_list$rho <- as.character(biome_list$rho)
  biome_list$rho[which(str_detect(biome_list$rho, "-1")==TRUE)] <- NA
  biome_list$rho <- as.numeric(biome_list$rho)
  biome_list$rho[which(biome_list$rho==-1)] <- NA
  biome_list$rho[which(biome_list$rho==1)] <- NA
  biome_list$z <- lapply(biome_list$rho, FisherZ)
  biome_list$v <- 1/(biome_list$n-3)
  biome_list <- as.data.table(biome_list)
  biome_list <- biome_list[, c("study_ID", "subject_id", "biome_ID") := tstrsplit(ID, "_")]
  biome_list <- as.data.frame(biome_list)
  biome_list$z <- as.numeric(biome_list$z) 
  biome_list$num_timepoint <- as.numeric(biome_list$num_timepoint)
  return(biome_list)
}

merged_gen.0001_final <- cleanup(merged_gen.0001)
merged_gen_no_art_final <- cleanup(merged_gen_no_art)

write.csv(merged_gen.0001_final, "merged_gen.0001_final.csv")
write.csv(merged_gen_no_art_final, "merged_gen_no_art_final.csv")
view(merged_gen_no_art_final)
```

I want to record and save all taxonomic information and associated c/e pairs
```{r}
#Unfortunately, since some rows were probably removed that were part of a family or genus with another row, I need to record all c/e pairs for all family and genera in the artifact and no-artifact datasets. This will get extracted from the max_results tables
extract_group_details <- function(study_list) {
  # Initialize an empty list to hold results
  result_list <- list()
  
  for (i in seq_along(study_list)) {
    study <- study_list[[i]]
    study_id <- names(study_list)[i]
    
    if (!"details" %in% names(study)) next
    
    details <- study$details
    
    # Skip if details is not a data.frame or data.table
    if (!is.data.frame(details)) next
    
    # Remove the first row if Group is empty string
    filtered_details <- details[details$Group != "", ]
    
    if (nrow(filtered_details) == 0) next
    
    # Add ID column
    filtered_details$ID <- study_id
    
    # Select only the desired columns
    selected <- filtered_details[, c("ID", "Group", "c", "e", "N", "NLL")]
    
    result_list[[length(result_list) + 1]] <- selected
  }
  
  # Combine all into a single data frame
  final_df <- do.call(rbind, result_list)
  return(final_df)
}

all_fam_.0001 <- extract_group_details(max_results.0001)
all_fam_no_art <- extract_group_details(max_results_no_art)
all_gen_.0001 <- extract_group_details(max_results_gen.0001)
all_gen_no_art <- extract_group_details(max_results_gen_no_art)

write.csv(all_fam_.0001, "fam_ce_values.0001.csv")
write.csv(all_fam_no_art, "fam_ce_values_no_art.csv")
write.csv(all_gen_.0001, "gen_ce_values.0001.csv")
write.csv(all_gen_no_art, "gen_ce_values_no_art.csv")
write.csv(removed_no_artifacts ,"all_artifact_taxa.csv")

OTU_tables_gen_fam_with_without_art <- list(data_.0001_clean, tables_no_artifacts)

write_rds(OTU_tables_gen_fam_with_without_art, "all_tables_art_and_no.rds")

```

Below this point is where I will add the code for analyzing the global metrics of community beta diversity once I clean those up.

FAVA!!!
```{r}
library(FAVA)

#FAVA is a normalized compositional variation statistic for things like OTU tables. It demands a matrix with samples as rows, taxa as columns, and can be weighted by timepoint if I include timepoint columns.
#First get the data:
OTU_tables <- read_rds("data_.0001_noChloroplast.rds")
included_names <- my_data.0001$ID
final_whole_data <- OTU_tables[included_names]
final_whole_data <- lapply(final_whole_data, function(df){
  df <- df[,-c(1:9)]
  return(df)
})
final_whole_data <- lapply(final_whole_data, t)
#convert to proportions
final_whole_data <- lapply(final_whole_data, function(df){
  df <- df[which(rowSums(df)>0),]
  df <- df / rowSums(df)
  return(df)
})
final_whole_data <- lapply(final_whole_data, function(df){
  df <- as.data.frame(df)
  df$time <- as.numeric(rownames(df))
  df <- df[,c(ncol(df), 1:(ncol(df)-1))]
  return(df)
})

##dataset is ready
fava_beans <- lapply(final_whole_data, function(df){
  bean <- fava(df, 
     K = ncol(df)-1,
     time = "time")
})
fava_beans <- do.call(rbind, fava_beans)
fava_df <- as.data.frame(fava_beans)
fava_df$ID <- rownames(fava_df)
fava_df$fava <- fava_df$V1
fava_df <- fava_df[,-1]
write.csv(fava_df, "fava_values.csv")

```
```{r}
#Troubleshooting
all_tables_half <- all_tables_art_and_no[[1]]

decimal_check <- function(table){
  columns <- colnames(table[,-c(1:9)])
  columns <- as.numeric(columns)
  dec <- which(columns < 1)
  print(dec)
}

decimal_check(all_tables_half$M6273_reindeer_lung)

lapply(all_tables_half,decimal_check)

```

