---
title: "Cleanish_metaanalysis_metaregression"
author: "Liam Nokes"
date: "`r Sys.Date()`"
output: html_document
---

Packages
```{r}
##Loading required packages
library(metafor)
library(clubSandwich)
library(orchaRd)
library(MuMIn)
library(mice)
library(metagear)
library(here)
library(DT)
library(GoodmanKruskal)
library(ggplot2)
library(readxl)
library(stringr)
library(tidyverse)
library(plotly)
library(cowplot)
library(grDevices)
library(grid)
library(gridGraphics)
library(pander)
library(formatR)
setwd("Studies_scripts")
```


Data
```{r}
my_data.0001 <- read.csv("final_biome_list_true.0001.csv")
my_data.0001$true_biome[which(my_data.0001$true_biome=="respiratory_oral")] <- "oral"

stability_data <- read.csv("stability_df.csv")
merged_with_slope <- read.csv("merged_metrics.csv")
merged_data.0001 <- inner_join(my_data.0001, merged_with_slope, by = join_by(ID==ID))

amount_of_families <- read.csv("distinct_families.csv")
merged_data.0001 <- inner_join(merged_data.0001, amount_of_families, by = join_by(ID==ID))

pc_slope_data <- read.csv("pc_slopes.csv")
merged_data.0001 <- inner_join(merged_data.0001, pc_slope_data, by = join_by(ID==ID))

fava_data <- read.csv("fava_values.csv")
merged_data.0001 <- inner_join(merged_data.0001, fava_data, by = join_by(ID == ID))

##now we have a merged_data.0001
#view(merged_data.0001)
#removing all of the old info
no_rho_no_slope <- merged_data.0001[,c(2,11,15:26,31:39, 47)]
no_rho_no_slope$pub_year_simple <- no_rho_no_slope$pub_year-mean(no_rho_no_slope$pub_year)

#Adding in all of the new info
family.0001 <- read.csv("merged_data_nobeta.0001.csv")
family_no_art <- read.csv("merged_data_nobeta_no_art.csv")
genus.0001 <- read.csv("merged_gen.0001_final.csv")
genus_no_art <- read.csv("merged_gen_no_art_final.csv")
#view(family.0001)
#merging with the old info
family.0001_complete <- inner_join(family.0001, no_rho_no_slope, by = join_by(ID==ID))
family.0001_complete <- family.0001_complete %>%
  mutate(study_ID = ifelse(study_ID == "MT5147", "M5147", study_ID))
family.0001_complete$ste <- family.0001_complete$ste^2

family_no_art_complete <- inner_join(family_no_art, no_rho_no_slope, by = join_by(ID==ID))
family_no_art_complete <- family_no_art_complete %>%
  mutate(study_ID = ifelse(study_ID == "MT5147", "M5147", study_ID))  
family_no_art_complete$ste <- family_no_art_complete$ste^2

genus.0001_complete <- inner_join(genus.0001, no_rho_no_slope, by = join_by(ID==ID))
genus.0001_complete <- genus.0001_complete %>%
  mutate(study_ID = ifelse(study_ID == "MT5147", "M5147", study_ID)) 
genus.0001_complete$ste <- genus.0001_complete$ste^2

genus_no_art_complete <- inner_join(genus_no_art, no_rho_no_slope, by = join_by(ID==ID))
genus_no_art_complete <- genus_no_art_complete %>%
  mutate(study_ID = ifelse(study_ID == "MT5147", "M5147", study_ID))
genus_no_art_complete$ste <- genus_no_art_complete$ste^2


#######These four datasets are the ones I will use for the meta-analysis^
datasets <- list(
  family.0001 = family.0001_complete,
  family.no_art = family_no_art_complete,
  genus.0001 = genus.0001_complete,
  genus.no_art = genus_no_art_complete
)

datasets_slim <- datasets
datasets_slim$family.0001 <- datasets_slim$family.0001[-which(datasets_slim$family.0001$distinct_families < 10),]
datasets_slim$family.no_art <- datasets_slim$family.no_art[-which(datasets_slim$family.no_art$distinct_families < 10),]
datasets_slim$genus.0001 <- datasets_slim$genus.0001[-which(datasets_slim$genus.0001$distinct_genera < 10),]
datasets_slim$genus.no_art <- datasets_slim$genus.no_art[-which(datasets_slim$genus.no_art$distinct_genera < 10),]

#counting participants and biomes
participants_biomes <- do.call(rbind,lapply(datasets_slim, function(df){
  participants <- length(unique(df$subject_id))
  biomes <- length(unique(df$ID))
  studies <- length(unique(df$study_ID))
  row <- data.frame("participants" = participants, "biomes" = biomes, "studies" = studies)
  return(row)
}))

###Adding these in case I add more analyses later
family_ce_info <- read.csv("fam_ce_values.0001.csv")
family_ce_no_art <- read.csv("fam_ce_values_no_art.csv")
genus_ce_info <- read.csv("gen_ce_values.0001.csv")
genus_ce_no_art <- read.csv("gen_ce_values_no_art.csv")
all_removed <- read.csv("all_artifact_taxa.csv")

```

1.1 Bias analysis
  a. visualize and address bias issues
  b. do "leave-one-out" plots to visualize sensitivity issues
```{r}
#Publication year bias (often a negative relationship between study year and effect size)
run_publication_year_bias_test <- function(data, effect_size_col, variance_col, level, no_art = FALSE) {
  # Subset data
  min_taxa_col <- ifelse(level == "family", "distinct_families", "distinct_genera")
  data_filtered <- data[data[[min_taxa_col]] > 9, ]
  
  # Model label
  art_status <- ifelse(no_art, "no_art", "0001")
  model_label <- paste0("mod_", ifelse(effect_size_col == "z", "ML_Z", "slope"),
                        "_pub_", substr(level, 1, 3), ".", art_status)
  
  # Try sparse = TRUE first
  cat("##########################\n")
  cat(model_label, "\n")
  
  try_sparse <- try({
    model <- rma.mv(yi = data_filtered[[effect_size_col]],
                    V = data_filtered[[variance_col]],
                    mods = ~ pub_year_simple,
                    random = list(~1 | study_ID,
                                  ~1 | subject_id,
                                  ~1 | biome_ID),
                    method = "REML",
                    data = data_filtered,
                    sparse = TRUE)
  }, silent = TRUE)

  if (inherits(try_sparse, "try-error")) {
    cat("Initial model failed to converge with sparse = TRUE. Retrying with sparse = FALSE...\n")
    model <- rma.mv(yi = data_filtered[[effect_size_col]],
                    V = data_filtered[[variance_col]],
                    mods = ~ pub_year_simple,
                    random = list(~1 | study_ID,
                                  ~1 | subject_id,
                                  ~1 | biome_ID),
                    method = "REML",
                    data = data_filtered,
                    sparse = FALSE)
  } else {
    model <- try_sparse
  }

  print(summary(model))
  return(model)
}


pub_models <- list()

for (name in names(datasets)) {
  dat <- datasets[[name]]
  level <- ifelse(grepl("genus", name), "genus", "family")
  no_art <- grepl("no_art", name)
  
  # RHO effect size
  pub_models[[paste0("ML_Z_", name)]] <- run_publication_year_bias_test(dat, effect_size_col = "z", variance_col = "v", level = level, no_art = no_art)
  
  # BETA effect size
  pub_models[[paste0("slope_", name)]] <- run_publication_year_bias_test(dat, effect_size_col = "slope", variance_col = "ste", level = level, no_art = no_art)
}

extract_model_summary <- function(model_list) {
  results <- lapply(names(model_list), function(name) {
    model <- model_list[[name]]

    coefs <- model$beta
    se <- model$se
    zval <- model$zval
    pval <- model$pval
    ci.lb <- model$ci.lb
    ci.ub <- model$ci.ub

    # Use rownames since beta is a matrix
    coef_names <- rownames(coefs)

    # Sanity check
    if (is.null(coef_names) || length(coef_names) != length(coefs)) {
      return(data.frame(
        model = name,
        term = NA,
        estimate = NA,
        se = NA,
        zval = NA,
        pval = NA,
        ci.lb = NA,
        ci.ub = NA,
        QM_stat = model$QM,
        QM_df = model$QMdf[1],
        QM_pval = model$QMp,
        QE_stat = model$QE,
        QE_df = model$QEdf,
        QE_pval = model$QEp,
        k = model$k,
        stringsAsFactors = FALSE
      ))
    }

    data.frame(
      model = rep(name, length(coefs)),
      term = coef_names,
      estimate = as.numeric(coefs),
      se = se,
      zval = zval,
      pval = pval,
      ci.lb = ci.lb,
      ci.ub = ci.ub,
      QM_stat = rep(model$QM, length(coefs)),
      QM_df = rep(model$QMdf[1], length(coefs)),
      QM_pval = rep(model$QMp, length(coefs)),
      QE_stat = rep(model$QE, length(coefs)),
      QE_df = rep(model$QEdf, length(coefs)),
      QE_pval = rep(model$QEp, length(coefs)),
      k = rep(model$k, length(coefs)),
      stringsAsFactors = FALSE
    )
  })

  do.call(rbind, results)
}

pub_model_table <- extract_model_summary(pub_models)
view(pub_models$ML_Z_family.0001$fit.stats)
################################################

#Detecting small study bias:
run_small_study_bias_test <- function(data, effect_size_col, variance_col, level, no_art = FALSE) {
  # Subset data
  min_taxa_col <- ifelse(level == "family", "distinct_families", "distinct_genera")
  data_filtered <- data[data[[min_taxa_col]] > 9, ]
  data_filtered$sqrt_v <- sqrt(data_filtered[[variance_col]])
  
  # Model label
  art_status <- ifelse(no_art, "no_art", "0001")
  model_label <- paste0("mod_", ifelse(effect_size_col == "z", "ML_Z", "slope"),
                        "_pub_", substr(level, 1, 3), ".", art_status)
  
  # Try sparse = TRUE first
  cat("##########################\n")
  cat(model_label, "\n")
  
  try_sparse <- try({
    model <- rma.mv(yi = data_filtered[[effect_size_col]],
                    V = data_filtered[[variance_col]],
                    mods = ~ sqrt_v,
                    random = list(~1 | study_ID,
                                  ~1 | subject_id,
                                  ~1 | biome_ID),
                    method = "REML",
                    data = data_filtered,
                    sparse = TRUE)
  }, silent = TRUE)

  if (inherits(try_sparse, "try-error")) {
    cat("Initial model failed to converge with sparse = TRUE. Retrying with sparse = FALSE...\n")
    model <- rma.mv(yi = data_filtered[[effect_size_col]],
                    V = data_filtered[[variance_col]],
                    mods = ~ sqrt_v,
                    random = list(~1 | study_ID,
                                  ~1 | subject_id,
                                  ~1 | biome_ID),
                    method = "REML",
                    data = data_filtered,
                    sparse = FALSE)
  } else {
    model <- try_sparse
  }

  print(summary(model))
  return(model)
}

small_models <- list()

for (name in names(datasets)) {
  dat <- datasets[[name]]
  level <- ifelse(grepl("genus", name), "genus", "family")
  no_art <- grepl("no_art", name)
  
  # RHO effect size
  small_models[[paste0("small_z_", name)]] <- run_small_study_bias_test(dat, effect_size_col = "z", variance_col = "v", level = level, no_art = no_art)
  
  # BETA effect size
  small_models[[paste0("small_slope_", name)]] <- run_small_study_bias_test(dat, effect_size_col = "slope", variance_col = "ste", level = level, no_art = no_art)
}

small_model_table <- extract_model_summary(small_models)


```

1.2 Sensitivity analysis (leave-one-out)
a. (leave one out for rho, biome level)
```{r}
# leave one out analysis
leave1out_ES <- list() # create a list to contain model estimates
est.func <- function(mod){
  df <- data.frame(est = mod$b, lower = mod$ci.lb, upper = mod$ci.ub)
  return(df)
}
family.0001_complete_slim <- family.0001_complete[-which(family.0001_complete$distinct_families<10),]
for(i in 1:length(family.0001_complete_slim$ID)){
  # create the data with one study removed at a time
  dat <- family.0001_complete_slim[-i,] # repeatedly run the multilevel model, leaving out one study at a time
  
  leave_mod <- rma.mv(yi = z, 
                               V = v, 
                               random = list(~1 | study_ID, # allows true effect sizes to vary among different primary studies - account for the between-study effect and quantify between-study heterogeneity;
                                             ~1 | subject_id,
                                             ~1 | biome_ID), 
                               method = "REML", 
                               test = "t",
                               data = dat,
                               sparse = TRUE) # model fitting
  leave1out_ES[[i]] <- est.func(leave_mod)}


names(leave1out_ES) <- family.0001_complete_slim$ID
leave1out_df <- do.call(rbind, lapply(leave1out_ES, function(df){
  row <- data.frame("est" = df[1], "lower" = df[2], "upper" = df[3])
  return(row)
}))

leave1out_df$Study <- rownames(leave1out_df)
leave1out_df$Study <- factor(leave1out_df$Study, levels = rev(rownames(leave1out_df)))

write.csv(leave1out_df, "leave_one_out.csv")


# Create the plot
ggplot(leave1out_df, aes(x = est, y = Study)) +
  geom_point() +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  xlab("Z (95% CI)") +
  ylab("Participant/biome left out") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 7),
    axis.text.x = element_text(size = 10),
    axis.title = element_text(size = 12)
  )
##SAVE AS 50/5
#The outlier is M1278_A_oral, but it isn't that much of an outlier (still within 95% confidence intervals)
#it appears the reason it is an outlier is not because it has a particularly high effect size
#but instead because it strongly affects the mean of its participant as well as it's study
#it's participant has 2 biomes sampled, and biome seems to have a profound effect here
#in the gut, the trade off is nonexistant but in the skin it is clearly present.

#doing study level now
leave1out_ST <- list()
for(i in 1:length(unique(family.0001_complete_slim$study_ID))){
  # create the data with one study removed at a time
  study_ids <- unique(family.0001_complete_slim$study_ID)
  dat <- family.0001_complete_slim[-which(family.0001_complete_slim$study_ID == study_ids[i]),] # repeatedly run the multilevel model, leaving out one study at a time
  print(study_ids[i])
  leave_mod <- rma.mv(yi = z, 
                      V = v, 
                      random = list(~1 | study_ID, # allows true effect sizes to vary among different primary studies - account for the between-study effect and quantify between-study heterogeneity;
                                    ~1 | subject_id,
                                    ~1 | biome_ID), 
                      method = "REML", 
                      test = "t",
                      data = dat,
                      sparse = TRUE) # model fitting
  print("fitted")
  leave1out_ST[[i]] <- est.func(leave_mod)}

names(leave1out_ST) <- unique(family.0001_complete_slim$study_ID)
leave1out_STdf <- do.call(rbind, lapply(leave1out_ST, function(df){
  row <- data.frame("est" = df[1], "lower" = df[2], "upper" = df[3])
  return(row)
}))

leave1out_STdf$Study <- rownames(leave1out_STdf)
write.csv(leave1out_STdf, "leave_one_study_out.csv")


#####################################NO ARTIFACTS
leave1out_ES_no_art <- list() # create a list to contain model estimates
family_no_art_complete_slim <- family_no_art_complete[-which(family_no_art_complete$distinct_families<10),]
for(i in 1:length(family_no_art_complete_slim$ID)){
  # create the data with one study removed at a time
  dat <- family_no_art_complete_slim[-i,] # repeatedly run the multilevel model, leaving out one study at a time
  print(family_no_art_complete_slim$ID[i])
  leave_mod <- rma.mv(yi = z, 
                               V = v, 
                               random = list(~1 | study_ID, # allows true effect sizes to vary among different primary studies - account for the between-study effect and quantify between-study heterogeneity;
                                             ~1 | subject_id,
                                             ~1 | biome_ID), 
                               method = "REML", 
                               test = "t",
                               data = dat,
                               sparse = TRUE) # model fitting
  leave1out_ES_no_art[[i]] <- est.func(leave_mod)}

names(leave1out_ES_no_art) <- family_no_art_complete_slim$ID
leave1out_df_no_art <- do.call(rbind, lapply(leave1out_ES_no_art, function(df){
  row <- data.frame("est" = df[1], "lower" = df[2], "upper" = df[3])
  return(row)
}))

leave1out_df_no_art$Study <- rownames(leave1out_df_no_art)
leave1out_df_no_art$Study <- factor(leave1out_df_no_art$Study, levels = rev(rownames(leave1out_df_no_art)))

write.csv(leave1out_df_no_art, "leave_one_out_no_art.csv")


# Create the plot
ggplot(leave1out_df_no_art, aes(x = est, y = Study)) +
  geom_point() +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  xlab("Z (95% CI)") +
  ylab("Participant/biome left out") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 7),
    axis.text.x = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

#doing study level now
leave1out_ST_no_art <- list()
for(i in 1:length(unique(family_no_art_complete_slim$study_ID))){
  # create the data with one study removed at a time
  study_ids <- unique(family_no_art_complete_slim$study_ID)
  dat <- family_no_art_complete_slim[-which(family_no_art_complete_slim$study_ID == study_ids[i]),] # repeatedly run the multilevel model, leaving out one study at a time
  print(study_ids[i])
  leave_mod <- rma.mv(yi = z, 
                      V = v, 
                      random = list(~1 | study_ID, # allows true effect sizes to vary among different primary studies - account for the between-study effect and quantify between-study heterogeneity;
                                    ~1 | subject_id,
                                    ~1 | biome_ID), 
                      method = "REML", 
                      test = "t",
                      data = dat,
                      sparse = TRUE) # model fitting
  print("fitted")
  leave1out_ST_no_art[[i]] <- est.func(leave_mod)}

names(leave1out_ST_no_art) <- unique(family_no_art_complete_slim$study_ID)
leave1out_STdf_no_art <- do.call(rbind, lapply(leave1out_ST_no_art, function(df){
  row <- data.frame("est" = df[1], "lower" = df[2], "upper" = df[3])
  return(row)
}))

leave1out_STdf_no_art$Study <- rownames(leave1out_STdf_no_art)
write.csv(leave1out_STdf_no_art, "leave_one_study_out_no_art.csv")



###################################################################
###################################################################
#Doing individual leave-one-out for the ones I haven't done yet (family with slopes, all genus-level)


leave1out_ES_no_art_gen <- list() # create a list to contain model estimates
genus_no_art_complete_slim <- genus_no_art_complete[-which(genus_no_art_complete$distinct_genera<10),]
for(i in 1:length(genus_no_art_complete_slim$ID)){
  # create the data with one study removed at a time
  dat <- genus_no_art_complete_slim[-i,] # repeatedly run the multilevel model, leaving out one study at a time
  print(genus_no_art_complete_slim$ID[i])
  leave_mod <- rma.mv(yi = z, 
                               V = v, 
                               random = list(~1 | study_ID, # allows true effect sizes to vary among different primary studies - account for the between-study effect and quantify between-study heterogeneity;
                                             ~1 | subject_id,
                                             ~1 | biome_ID), 
                               method = "REML", 
                               test = "t",
                               data = dat,
                               sparse = TRUE) # model fitting
  leave1out_ES_no_art_gen[[i]] <- est.func(leave_mod)}

names(leave1out_ES_no_art_gen) <- genus_no_art_complete_slim$ID
leave1out_df_no_art_gen <- do.call(rbind, lapply(leave1out_ES_no_art_gen, function(df){
  row <- data.frame("est" = df[1], "lower" = df[2], "upper" = df[3])
  return(row)
}))

leave1out_df_no_art_gen$Study <- rownames(leave1out_df_no_art_gen)
leave1out_df_no_art_gen$Study <- factor(leave1out_df_no_art_gen$Study, levels = rev(rownames(leave1out_df_no_art_gen)))

write.csv(leave1out_df_no_art_gen, "leave_one_out_no_art_gen.csv")

###Genus with artifacts (rho)
leave1out_ES.0001_gen <- list() # create a list to contain model estimates
genus.0001_complete_slim <- genus.0001_complete[-which(genus.0001_complete$distinct_genera<10),]
for(i in 1:length(genus.0001_complete_slim$ID)){
  # create the data with one study removed at a time
  dat <- genus.0001_complete_slim[-i,] # repeatedly run the multilevel model, leaving out one study at a time
  print(genus.0001_complete_slim$ID[i])
  leave_mod <- rma.mv(yi = z, 
                               V = v, 
                               random = list(~1 | study_ID, # allows true effect sizes to vary among different primary studies - account for the between-study effect and quantify between-study heterogeneity;
                                             ~1 | subject_id,
                                             ~1 | biome_ID), 
                               method = "REML", 
                               test = "t",
                               data = dat,
                               sparse = TRUE) # model fitting
  leave1out_ES.0001_gen[[i]] <- est.func(leave_mod)}

names(leave1out_ES.0001_gen) <- genus.0001_complete_slim$ID
leave1out_df.0001_gen <- do.call(rbind, lapply(leave1out_ES.0001_gen, function(df){
  row <- data.frame("est" = df[1], "lower" = df[2], "upper" = df[3])
  return(row)
}))

leave1out_df.0001_gen$Study <- rownames(leave1out_df.0001_gen)
leave1out_df.0001_gen$Study <- factor(leave1out_df.0001_gen$Study, levels = rev(rownames(leave1out_df.0001_gen)))

write.csv(leave1out_df.0001_gen, "leave_one_out.0001_gen.csv")

##################################################################################################################################################################################



```


1.2.b. leave one out by slope



```{r}
###family level with artifacts
leave1out_slope_ES <- list()
for(i in 1:length(family.0001_complete_slim$ID)){
  # create the data with one study removed at a time
  dat <- family.0001_complete_slim[-i,] # repeatedly run the multilevel model, leaving out one study at a time
  print(family.0001_complete_slim$ID[i])
  leave_mod <- rma.mv(yi = slope, 
                               V = ste, 
                               random = list(~1 | study_ID, # allows true effect sizes to vary among different primary studies - account for the between-study effect and quantify between-study heterogeneity;
                                             ~1 | subject_id,
                                             ~1 | biome_ID), 
                               method = "REML", 
                               test = "t",
                               data = dat,
                               sparse = TRUE) # model fitting
  leave1out_slope_ES[[i]] <- est.func(leave_mod)}


names(leave1out_slope_ES) <- family.0001_complete_slim$ID
leave1out_slope_df <- do.call(rbind, lapply(leave1out_slope_ES, function(df){
  row <- data.frame("est" = df[1], "lower" = df[2], "upper" = df[3])
  return(row)
}))

leave1out_slope_df$Study <- rownames(leave1out_slope_df)
leave1out_slope_df$Study <- factor(leave1out_slope_df$Study, levels = rev(rownames(leave1out_slope_df)))

write.csv(leave1out_slope_df, "leave_one_out_slope.csv")

####Family level no artifacts
leave1out_ES_slope_no_art <- list()

for (i in seq_along(family_no_art_complete_slim$ID)) {
  # Leave one study out
  dat <- family_no_art_complete_slim[-i, ]
  study_out <- family_no_art_complete_slim$ID[i]
  print(study_out)

  # Try sparse = TRUE first
  leave_mod <- tryCatch({
    rma.mv(
      yi = slope, 
      V = ste, 
      random = list(~1 | study_ID, ~1 | subject_id, ~1 | biome_ID), 
      method = "REML", 
      test = "t", 
      data = dat,
      sparse = TRUE
    )
  }, error = function(e) {
    message(paste("Sparse model failed for", study_out, "- trying sparse = FALSE"))
    # Try again without sparse
    tryCatch({
      rma.mv(
        yi = slope, 
        V = ste, 
        random = list(~1 | study_ID, ~1 | subject_id, ~1 | biome_ID), 
        method = "REML", 
        test = "t", 
        data = dat,
        sparse = FALSE
      )
    }, error = function(e2) {
      message(paste("Model failed completely for", study_out))
      return(NULL)
    })
  })

  leave1out_ES_slope_no_art[[i]] <- if (!is.null(leave_mod)) est.func(leave_mod) else NA
}

names(leave1out_ES_slope_no_art) <- family_no_art_complete_slim$ID
leave1out_df_slope_no_art <- do.call(rbind, lapply(leave1out_ES_slope_no_art, function(df){
  row <- data.frame("est" = df[1], "lower" = df[2], "upper" = df[3])
  return(row)
}))

leave1out_df_slope_no_art$Study <- rownames(leave1out_df_slope_no_art)
leave1out_df_slope_no_art$Study <- factor(leave1out_df_slope_no_art$Study, levels = rev(rownames(leave1out_df_slope_no_art)))

write.csv(leave1out_df_slope_no_art, "leave_one_out_slope_no_art.csv")

###########################Genus with artifacts
leave1out_ES.0001_slope_gen <- list() # create a list to contain model estimates
for(i in 1:length(genus.0001_complete_slim$ID)){
  # create the data with one study removed at a time
  dat <- genus.0001_complete_slim[-i,] # repeatedly run the multilevel model, leaving out one study at a time
  print(genus.0001_complete_slim$ID[i])
  leave_mod <- rma.mv(yi = slope, 
                               V = ste, 
                               random = list(~1 | study_ID, # allows true effect sizes to vary among different primary studies - account for the between-study effect and quantify between-study heterogeneity;
                                             ~1 | subject_id,
                                             ~1 | biome_ID), 
                               method = "REML", 
                               test = "t",
                               data = dat,
                               sparse = TRUE) # model fitting
  leave1out_ES.0001_slope_gen[[i]] <- est.func(leave_mod)}

names(leave1out_ES.0001_slope_gen) <- genus.0001_complete_slim$ID
leave1out_df.0001_slope_gen <- do.call(rbind, lapply(leave1out_ES.0001_slope_gen, function(df){
  row <- data.frame("est" = df[1], "lower" = df[2], "upper" = df[3])
  return(row)
}))

leave1out_df.0001_slope_gen$Study <- rownames(leave1out_df.0001_slope_gen)
leave1out_df.0001_slope_gen$Study <- factor(leave1out_df.0001_slope_gen$Study, levels = rev(rownames(leave1out_df.0001_slope_gen)))

write.csv(leave1out_df.0001_slope_gen, "leave_one_out.0001_slope_gen.csv")

#####################################################################
####Genus no artifacts
leave1out_ES_slope_no_art_gen <- list()

for (i in seq_along(genus_no_art_complete_slim$ID)) {
  # Leave one study out
  dat <- genus_no_art_complete_slim[-i, ]
  study_out <- genus_no_art_complete_slim$ID[i]
  print(study_out)

  # Try sparse = TRUE first
  leave_mod <- tryCatch({
    rma.mv(
      yi = slope, 
      V = ste, 
      random = list(~1 | study_ID, ~1 | subject_id, ~1 | biome_ID), 
      method = "REML", 
      test = "t", 
      data = dat,
      sparse = TRUE
    )
  }, error = function(e) {
    message(paste("Sparse model failed for", study_out, "- trying sparse = FALSE"))
    # Try again without sparse
    tryCatch({
      rma.mv(
        yi = slope, 
        V = ste, 
        random = list(~1 | study_ID, ~1 | subject_id, ~1 | biome_ID), 
        method = "REML", 
        test = "t", 
        data = dat,
        sparse = FALSE
      )
    }, error = function(e2) {
      message(paste("Model failed completely for", study_out))
      return(NULL)
    })
  })

  leave1out_ES_slope_no_art_gen[[i]] <- if (!is.null(leave_mod)) est.func(leave_mod) else NA
}

names(leave1out_ES_slope_no_art_gen) <- genus_no_art_complete_slim$ID
leave1out_df_slope_no_art_gen <- do.call(rbind, lapply(leave1out_ES_slope_no_art_gen, function(df){
  row <- data.frame("est" = df[1], "lower" = df[2], "upper" = df[3])
  return(row)
}))

leave1out_df_slope_no_art_gen$Study <- rownames(leave1out_df_slope_no_art_gen)
leave1out_df_slope_no_art_gen$Study <- factor(leave1out_df_slope_no_art_gen$Study, levels = rev(rownames(leave1out_df_slope_no_art_gen)))

write.csv(leave1out_df_slope_no_art_gen, "leave_one_out_slope_no_art_gen.csv")

######################################################################
#########################ALLL STUDY LEVEL SENSITIVITY ANALYSIS########
######################################################################

run_leave_one_out_analysis <- function(data, level = "family",
                                       effect_size_col = "z",
                                       variance_col = "v",
                                       est.func,
                                       sparse = TRUE) {
  # Filter by minimum taxa
  min_taxa_col <- ifelse(level == "family", "distinct_families", "distinct_genera")
  data_slim <- data[data[[min_taxa_col]] > 9, ]

  study_ids <- unique(data_slim$study_ID)
  leave1out_results <- vector("list", length(study_ids))
  names(leave1out_results) <- study_ids

  for (i in seq_along(study_ids)) {
    sid <- study_ids[i]
    cat("Leaving out study:", sid, "\n")

    dat <- data_slim[data_slim$study_ID != sid, ]

    leave_mod <- tryCatch({
      rma.mv(yi = dat[[effect_size_col]],
             V = dat[[variance_col]],
             random = list(~1 | study_ID,
                           ~1 | subject_id,
                           ~1 | biome_ID),
             method = "REML",
             test = "t",
             data = dat,
             sparse = sparse)
    }, error = function(e) {
      warning(sprintf("Model failed for study: %s", sid))
      return(NULL)
    })

    if (!is.null(leave_mod)) {
      leave1out_results[[i]] <- est.func(leave_mod)
    } else {
      leave1out_results[[i]] <- c(NA, NA, NA)
    }
  }

  leave1out_df <- do.call(rbind, lapply(leave1out_results, function(res) {
    data.frame(est = res[1], lower = res[2], upper = res[3])
  }))
  leave1out_df$Study <- names(leave1out_results)
  rownames(leave1out_df) <- NULL

  return(leave1out_df)
}

leave1out_all_studies <- list()

for (name in names(datasets)) {
  dat <- datasets[[name]]
  level <- ifelse(grepl("genus", name), "genus", "family")

  # Rho (z)
  res_z <- run_leave_one_out_analysis(
    data = dat,
    level = level,
    effect_size_col = "z",
    variance_col = "v",
    est.func = est.func
  )
  res_z$model <- name
  res_z$effect_type <- "rho"

  # Slope (beta)
  res_slope <- run_leave_one_out_analysis(
    data = dat,
    level = level,
    effect_size_col = "slope",
    variance_col = "ste",
    est.func = est.func
  )
  res_slope$model <- name
  res_slope$effect_type <- "slope"

  # Store both
  leave1out_all_studies[[paste0(name, "_z")]] <- res_z
  leave1out_all_studies[[paste0(name, "_slope")]] <- res_slope
}

leave1out_all_studies

```

1.3 figures for sensitivity analysis
```{r}
leave_one_out.0001_fam_rho <- read.csv("leave_one_out.csv")
leave_one_out.0001_gen_rho <- read.csv("leave_one_out.0001_gen.csv")
leave_one_out_no_art_fam_rho <- read.csv("leave_one_out_no_art.csv")
leave_one_out_no_art_gen_rho <- read.csv("leave_one_out_no_art_gen.csv")
leave_one_out.0001_fam_slope <- read.csv("leave_one_out_slope.csv")
leave_one_out.0001_gen_slope <- read.csv("leave_one_out.0001_slope_gen.csv")
leave_one_out_no_art_fam_slope <- read.csv("leave_one_out_slope_no_art.csv")
leave_one_out_no_art_gen_slope <- read.csv("leave_one_out_slope_no_art_gen.csv")

plot_list <- list(
  leave1 = ggplot(leave_one_out.0001_fam_rho, aes(x = est, y = Study)) + 
    geom_point() + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    xlab("Z (95% CI)") + ylab("Participant/biome left out") + ggtitle("Leave one out - Family") +
    theme_minimal() + theme(axis.text.y = element_text(size = 7), axis.text.x = element_text(size = 10), axis.title = element_text(size = 12)),

  leave2 = ggplot(leave_one_out_no_art_fam_rho, aes(x = est, y = Study)) +
    geom_point() + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    xlab("Z (95% CI)") + ylab("Participant/biome left out") + ggtitle("Leave one out - Family (no artifacts)") +
    theme_minimal() + theme(axis.text.y = element_text(size = 7), axis.text.x = element_text(size = 10), axis.title = element_text(size = 12)),

  leave3 = ggplot(leave_one_out.0001_gen_rho, aes(x = est, y = Study)) +
    geom_point() + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    xlab("Z (95% CI)") + ylab("Participant/biome left out") + ggtitle("Leave one out - Genus") +
    theme_minimal() + theme(axis.text.y = element_text(size = 7), axis.text.x = element_text(size = 10), axis.title = element_text(size = 12)),

  leave4 = ggplot(leave_one_out_no_art_gen_rho, aes(x = est, y = Study)) +
    geom_point() + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    xlab("Z (95% CI)") + ylab("Participant/biome left out") + ggtitle("Leave one out - Genus (no artifacts)") +
    theme_minimal() + theme(axis.text.y = element_text(size = 7), axis.text.x = element_text(size = 10), axis.title = element_text(size = 12)),

  leave5 = ggplot(leave_one_out.0001_fam_slope, aes(x = est, y = Study)) +
    geom_point() + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    xlab("Beta (95% CI)") + ylab("Participant/biome left out") + ggtitle("Leave one out - Family") +
    theme_minimal() + theme(axis.text.y = element_text(size = 7), axis.text.x = element_text(size = 10), axis.title = element_text(size = 12)),

  leave6 = ggplot(leave_one_out_no_art_fam_slope, aes(x = est, y = Study)) +
    geom_point() + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    xlab("Beta (95% CI)") + ylab("Participant/biome left out") + ggtitle("Leave one out - Family (no artifacts)") +
    theme_minimal() + theme(axis.text.y = element_text(size = 7), axis.text.x = element_text(size = 10), axis.title = element_text(size = 12)),

  leave7 = ggplot(leave_one_out.0001_gen_slope, aes(x = est, y = Study)) +
    geom_point() + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    xlab("Beta (95% CI)") + ylab("Participant/biome left out") + ggtitle("Leave one out - Genus") +
    theme_minimal() + theme(axis.text.y = element_text(size = 7), axis.text.x = element_text(size = 10), axis.title = element_text(size = 12)),

  leave8 = ggplot(leave_one_out_no_art_gen_slope, aes(x = est, y = Study)) +
    geom_point() + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    xlab("Beta (95% CI)") + ylab("Participant/biome left out") + ggtitle("Leave one out - Genus (no artifacts)") +
    theme_minimal() + theme(axis.text.y = element_text(size = 7), axis.text.x = element_text(size = 10), axis.title = element_text(size = 12)),

  # All studies
  leave9 = ggplot(leave1out_all_studies$family.0001_z, aes(x = est, y = Study)) + 
    geom_point() + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed") + xlab("Z (95% CI)") + ylab("Study left out") +
    ggtitle("Leave one study out - Family") + theme_minimal() + 
    theme(axis.text.y = element_text(size = 7), axis.text.x = element_text(size = 10), axis.title = element_text(size = 12)),

  leave10 = ggplot(leave1out_all_studies$family.no_art_z, aes(x = est, y = Study)) +
    geom_point() + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed") + xlab("Z (95% CI)") + ylab("Study left out") +
    ggtitle("Leave one study out - Family (no artifacts)") + theme_minimal() +
    theme(axis.text.y = element_text(size = 7), axis.text.x = element_text(size = 10), axis.title = element_text(size = 12)),

  leave11 = ggplot(leave1out_all_studies$genus.0001_z, aes(x = est, y = Study)) +
    geom_point() + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed") + xlab("Z (95% CI)") + ylab("Study left out") +
    ggtitle("Leave one study out - Genus") + theme_minimal() +
    theme(axis.text.y = element_text(size = 7), axis.text.x = element_text(size = 10), axis.title = element_text(size = 12)),

  leave12 = ggplot(leave1out_all_studies$genus.no_art_z, aes(x = est, y = Study)) +
    geom_point() + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed") + xlab("Z (95% CI)") + ylab("Study left out") +
    ggtitle("Leave one study out - Genus (no artifacts)") + theme_minimal() +
    theme(axis.text.y = element_text(size = 7), axis.text.x = element_text(size = 10), axis.title = element_text(size = 12)),

  leave13 = ggplot(leave1out_all_studies$family.0001_slope, aes(x = est, y = Study)) +
    geom_point() + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed") + xlab("Beta (95% CI)") + ylab("Study left out") +
    ggtitle("Leave one study out - Family") + theme_minimal() +
    theme(axis.text.y = element_text(size = 7), axis.text.x = element_text(size = 10), axis.title = element_text(size = 12)),

  leave14 = ggplot(leave1out_all_studies$family.no_art_slope, aes(x = est, y = Study)) +
    geom_point() + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed") + xlab("Beta (95% CI)") + ylab("Study left out") +
    ggtitle("Leave one study out - Family (no artifacts)") + theme_minimal() +
    theme(axis.text.y = element_text(size = 7), axis.text.x = element_text(size = 10), axis.title = element_text(size = 12)),

  leave15 = ggplot(leave1out_all_studies$genus.0001_slope, aes(x = est, y = Study)) +
    geom_point() + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed") + xlab("Beta (95% CI)") + ylab("Study left out") +
    ggtitle("Leave one study out - Genus") + theme_minimal() +
    theme(axis.text.y = element_text(size = 7), axis.text.x = element_text(size = 10), axis.title = element_text(size = 12)),

  leave16 = ggplot(leave1out_all_studies$genus.no_art_slope, aes(x = est, y = Study)) +
    geom_point() + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed") + xlab("Beta (95% CI)") + ylab("Study left out") +
    ggtitle("Leave one study out - Genus (no artifacts)") + theme_minimal() +
    theme(axis.text.y = element_text(size = 7), axis.text.x = element_text(size = 10), axis.title = element_text(size = 12))
)

# Define file names
file_names <- paste0("Studies_scripts/figures/leaveout_plot_", seq_along(plot_list), ".pdf")

# Save all plots
for (i in seq_along(plot_list)) {
  dims <- if (i <= 8) c(5, 50) else c(5, 4)
  ggsave(file_names[i], plot = plot_list[[i]], width = dims[1], height = dims[2], units = "in", limitsize = FALSE)
}

```


2. get "global means", make forest plots (slope and rho)
```{r}
run_meta_model <- function(data, level = "family", effect_size_col = "z", variance_col = "v") {
  # Filter based on minimum taxa
  min_taxa_col <- ifelse(level == "family", "distinct_families", "distinct_genera")
  data_filtered <- data[data[[min_taxa_col]] > 9, ]

  # Fit multilevel model
  model <- rma.mv(
    yi = data_filtered[[effect_size_col]],
    V = data_filtered[[variance_col]],
    random = list(
      ~1 | study_ID,
      ~1 | subject_id,
      ~1 | biome_ID
    ),
    method = "REML",
    data = data_filtered,
    sparse = TRUE
  )

  return(model)
}

meta_models <- list()

for (name in names(datasets)) {
  dat <- datasets[[name]]
  level <- ifelse(grepl("genus", name), "genus", "family")

  # Rho model
  model_rho <- run_meta_model(dat, level = level, effect_size_col = "z", variance_col = "v")
  model_name_rho <- paste0("mod_rho_", name)
  meta_models[[model_name_rho]] <- model_rho

  # Slope model
  model_slope <- run_meta_model(dat, level = level, effect_size_col = "slope", variance_col = "ste")
  model_name_slope <- paste0("mod_slope_", name)
  meta_models[[model_name_slope]] <- model_slope
}

meta_model_summary <- extract_model_summary(meta_models)
write.csv(meta_model_summary, "meta_means.csv")

for (name in names(meta_models)) {
  pdf(paste0("Studies_scripts/figures/forest_", name, ".pdf"), width = 7, height = 5)
  forest(meta_models[[name]], xlab = ifelse(grepl("rho", name), "Effect size Z", "Effect size Beta"))
  dev.off()
}

```


3. Regress by biome, make orchard plots
```{r}
run_meta_model_biome <- function(data, level = "family", effect_size_col = "z", variance_col = "v") {
  # Filter based on distinct taxa
  min_taxa_col <- ifelse(level == "family", "distinct_families", "distinct_genera")
  data_filtered <- data[data[[min_taxa_col]] > 9, ]

  # Fit model with biome moderator (no intercept)
  model <- rma.mv(
    yi = data_filtered[[effect_size_col]],
    V = data_filtered[[variance_col]],
    mods = ~ true_biome - 1,
    random = list(
      ~1 | study_ID,
      ~1 | subject_id,
      ~1 | biome_ID
    ),
    method = "REML",
    test = "t",
    data = data_filtered,
    sparse = TRUE
  )

  return(model)
}

biome_models <- list()

for (name in names(datasets)) {
  dat <- datasets[[name]]
  level <- ifelse(grepl("genus", name), "genus", "family")

  # Rho (z)
  model_rho <- run_meta_model_biome(dat, level = level, effect_size_col = "z", variance_col = "v")
  biome_models[[paste0("biome_rho_", name)]] <- model_rho

  # Slope (beta)
  model_slope <- run_meta_model_biome(dat, level = level, effect_size_col = "slope", variance_col = "ste")
  biome_models[[paste0("biome_slope_", name)]] <- model_slope
}


biome_summary_table <- extract_model_summary(biome_models)


orchard_plot(biome_models$biome_rho_family.0001, 
             mod = "true_biome", 
             xlab = "Effect size Z",
             group = "study_ID",  k = TRUE, g = FALSE, trunk.size = 1.5,) + 
  scale_x_discrete(labels = c("gut","oral","respiratory","skin","vaginal"))

# Define custom biome labels
biome_labels <- c("gut", "oral", "resp.", "skin", "vaginal")

# Optional: clean labels for filenames
clean_filename <- function(name) {
  gsub("[^a-zA-Z0-9]", "_", name)
}

biome_models <- biome_models[c(3,4,7,8)]

# Loop over each model and save orchard plots
for (name in names(biome_models)) {
  model <- biome_models[[name]]

  # Set x-axis label based on effect type
  xlab_text <- if (grepl("rho", name)) {
    "Effect size Z"
  } else {
    "Effect size Beta"
  }

  # Build the plot
  p <- orchard_plot(model,
                    mod = "true_biome",
                    group = "study_ID",
                    xlab = xlab_text,
                    k = TRUE, g = FALSE,
                    trunk.size = 1.5) +
       scale_x_discrete(labels = biome_labels) +
       ggtitle(paste("Orchard plot:", name)) +
       theme_minimal()

  # Save to PDF
  ggsave(
    filename = paste0("Studies_scripts/figures/orchard_", clean_filename(name), ".pdf"),
    plot = p,
    width = 7, height = 5, units = "in"
  )
}

plot_list_biomes <- lapply(names(biome_models), function(name) {
  model <- biome_models[[name]]
  
  # Determine effect type and level
  effect <- if (str_detect(name, "slope")) "Beta" else "Z"
  level <- if (str_detect(name, "genus")) "Genus" else "Family"
  artifact <- if (str_detect(name, "no_art")) "No Artifacts" else "With Artifacts"
  
  # Create label
  plot_title <- paste(level, "-", effect, "-", artifact)
  
  # Make orchard plot
  p <- orchard_plot(
  model,
  mod = "true_biome",
  xlab = paste("Effect size", effect),
  group = "study_ID",
  k = TRUE,
  g = FALSE,
  trunk.size = 1.5
) +
  scale_x_discrete(labels = biome_labels)
  
  return(p)
})

# Combine with patchwork
library(patchwork)
combined_plot_biome <- wrap_plots(plotlist = plot_list_biomes, ncol = 2)
combined_plot_biome


```


4. Regress by beta diversity (richness BDtotal, BDprop, turnover)
```{r}
############################AUTOMATED
run_meta_model_diversity <- function(data, moderator, level = "family", effect_size_col = "z", variance_col = "v") {
  # Filter for sufficient taxa
  min_taxa_col <- ifelse(level == "family", "distinct_families", "distinct_genera")
  data_filtered <- data[data[[min_taxa_col]] > 9, ]

  # Fit model with given moderator
  formula_mod <- as.formula(paste("~", moderator))
  model <- rma.mv(
    yi = data_filtered[[effect_size_col]],
    V = data_filtered[[variance_col]],
    mods = formula_mod,
    random = list(~1 | study_ID, ~1 | subject_id, ~1 | biome_ID),
    method = "REML",
    test = "t",
    data = data_filtered,
    sparse = TRUE
  )
  return(model)
}

# List of moderators and datasets
moderators <- list(
  richness_log = "log(richness)",
  BDtotal = "BDtotal",
  rate_change = "rate_change",
  average_turnover = "turnover"
)

diversity_models <- list()

for (name in names(datasets)) {
  dat <- datasets[[name]]
  level <- ifelse(grepl("genus", name), "genus", "family")

  # Filter for distinct taxa > 9
  min_taxa_col <- ifelse(level == "family", "distinct_families", "distinct_genera")
  dat <- dat[dat[[min_taxa_col]] > 9, ]

  # Add log-transformed diversity measures
  dat$richness_log <- log(dat$richness)
  dat$BDtotal_log <- log(dat$BDtotal)
  dat$turnover_log <- log(dat$average_turnover)

  for (mod_name in names(moderators)) {
    # Rho (z)
    model_z <- run_meta_model_diversity(dat, moderator = mod_name, level = level, effect_size_col = "z", variance_col = "v")
    diversity_models[[paste0("diversity_", mod_name, "_z_", name)]] <- model_z

    # Slope (beta)
    model_slope <- run_meta_model_diversity(dat, moderator = mod_name, level = level, effect_size_col = "slope", variance_col = "ste")
    diversity_models[[paste0("diversity_", mod_name, "_slope_", name)]] <- model_slope
  }
}

# Save summaries if needed
diversity_summary <- extract_model_summary(diversity_models)

#Plotting
# Plotting loop for bubble plots
for (model_name in names(diversity_models)) {
  model <- diversity_models[[model_name]]

  mod_cols <- setdiff(colnames(model$X), "(Intercept)")
  if (length(mod_cols) == 0) {
    warning(paste("Skipping", model_name, "- no valid moderator found."))
    next
  }
  mod_term <- mod_cols[1]
  if (mod_term == "intrcpt" && ncol(model$X) > 1) mod_term <- colnames(model$X)[2]

  xlab_text <- case_when(
    grepl("richness_log", model_name) ~ "Richness (log)",
    grepl("BDtotal", model_name) ~ "Total community variance",
    grepl("rate_change", model_name) ~ "Time lag slope",
    grepl("turnover_log", model_name) ~ "Turnover (log)",
    TRUE ~ mod_term
  )

  p <- bubble_plot(model,
                   mod = mod_term,
                   xlab = xlab_text,
                   ylab = ifelse(grepl("slope", model_name), "Effect size Beta", "Effect size Z"),
                   group = "study_ID",
                   k = TRUE, g = TRUE,
                   legend.pos = "top.left") +
       theme(
         axis.text.x = element_text(size = 12, colour = "black"),
         axis.text.y = element_text(size = 12, colour = "black"),
         axis.title.x = element_text(size = 12, colour = "black"),
         plot.title = element_text(size = 12, colour = "black"))

  ggsave(
    filename = paste0("Studies_scripts/figures/bubble_", clean_filename(model_name), ".pdf"),
    plot = p,
    width = 6, height = 5, units = "in"
  )
}


library(patchwork)
clean_filename_1 <- function(x) gsub("[^[:alnum:]_]", "_", x)


facet_bubble_plots_by_moderator <- function(diversity_models, out_dir = "Studies_scripts/figures") {

  short_labels <- c(
    "family.0001" = "family",
    "family.no_art" = "family (no artifacts)",
    "genus.0001" = "genus",
    "genus.no_art" = "genus (no artifacts)"
  )

  # Extract moderator and effect type from model name
  extract_info <- function(name) {
    parts <- str_split(name, "_")[[1]]
    list(
      effect = ifelse("slope" %in% parts, "slope", "z"),
      moderator = parts[2],
      dataset = sub(".*_(family.*|genus.*)$", "\\1", name),
      full_name = name
    )
  }

  model_info <- lapply(names(diversity_models), extract_info) %>% bind_rows()

  model_info$label <- short_labels[model_info$dataset]

  # Group by moderator and effect type
  grouped <- model_info %>% group_by(moderator, effect) %>% group_split()

  for (group in grouped) {
    mod <- unique(group$moderator)
    eff <- unique(group$effect)

    plots <- list()

    for (i in seq_len(nrow(group))) {
      name <- group$full_name[i]
      model <- diversity_models[[name]]
      label <- group$label[i]

      mod_cols <- setdiff(colnames(model$X), "(Intercept)")
      if (length(mod_cols) == 0) next
      mod_term <- mod_cols[1]
      if (mod_term == "intrcpt" && ncol(model$X) > 1) mod_term <- colnames(model$X)[2]

      xlab_text <- case_when(
        grepl("log_fava", name) ~ "FAVA (log)",
        grepl("sqrt_freq", name) ~ "Inverse sampling frequency (sqrt)",
        grepl("rate_change", name) ~ "Time lag slope",
        grepl("average_turnover", name) ~ "Average turnover",
        grepl("pub_year", name) ~ "Mean-centered publication year",
        TRUE ~ mod_term
      )

      ylab_text <- ifelse(eff == "slope", "Effect size Beta", "Effect size Z")

      p <- bubble_plot(model,
                       mod = mod_term,
                       xlab = xlab_text,
                       ylab = ylab_text,
                       group = "study_ID",
                       k = FALSE, g = FALSE,
                       legend.pos = "top.left") +
        ggtitle(label) +
        theme(
          axis.text.x = element_text(size = 10, colour = "black"),
          axis.text.y = element_text(size = 10, colour = "black"),
          axis.title.x = element_text(size = 11, colour = "black"),
          axis.title.y = element_text(size = 11, colour = "black"),
          plot.title = element_text(size = 12, face = "bold", hjust = 0.5)
        )

      plots[[length(plots) + 1]] <- p
    }

    if (length(plots) > 0) {
      # Facet wrap in groups of 4
      plot_chunks <- split(plots, ceiling(seq_along(plots) / 4))
      for (j in seq_along(plot_chunks)) {
        fig <- wrap_plots(plot_chunks[[j]], ncol = 2)

        filename <- paste0(
          out_dir, "/facet_", mod, "_", eff, "_group", j, ".pdf"
        )

        ggsave(filename, fig, width = 10, height = 10, units = "in")
        message("Saved: ", filename)
      }
    }
  }
}

#run in the console
facet_bubble_plots_by_moderator(diversity_models)


```


5. Regress by antibiotics, probiotics, disease, location, infant?
```{r}
antibiotic_models <- list()

for (name in names(datasets)) {
  dat <- datasets[[name]]
  level <- ifelse(grepl("genus", name), "genus", "family")

  # Filter for distinct taxa > 9
  min_taxa_col <- ifelse(level == "family", "distinct_families", "distinct_genera")
  dat <- dat[dat[[min_taxa_col]] > 9, ]

  # Rho (z)
  model_z <- rma.mv(
    yi = z,
    V = v,
    mods = ~ development + disease + sustained_probiotic + sustained_antibiotic,
    random = list(~1 | study_ID, ~1 | subject_id, ~1 | biome_ID),
    method = "REML",
    test = "t",
    data = dat,
    sparse = TRUE
  )
  antibiotic_models[[paste0("antibiotic_z_", name)]] <- model_z

  # Slope (beta)
  model_slope <- rma.mv(
    yi = slope,
    V = ste,
    mods = ~ development + disease + sustained_probiotic + sustained_antibiotic,
    random = list(~1 | study_ID, ~1 | subject_id, ~1 | biome_ID),
    method = "REML",
    test = "t",
    data = dat,
    sparse = TRUE
  )
  antibiotic_models[[paste0("antibiotic_slope_", name)]] <- model_slope
}

antibiotic_summary <- extract_model_summary(antibiotic_models)
write.csv(antibiotic_summary, "Studies_scripts/antibiotic_model_summary.csv", row.names = FALSE)

# Bubble plots for each moderator in the antibiotic models
for (model_name in names(antibiotic_models)) {
  model <- antibiotic_models[[model_name]]

  mod_terms <- setdiff(colnames(model$X), "(Intercept)")
  for (mod_term in mod_terms) {
    p <- bubble_plot(model,
                     mod = mod_term,
                     xlab = mod_term,
                     ylab = ifelse(grepl("slope", model_name), "Effect size Beta", "Effect size Z"),
                     group = "study_ID",
                     k = TRUE, g = TRUE,
                     legend.pos = "top.left") +
         ggtitle(paste("Bubble plot:", mod_term, "in", model_name)) +
         theme(
           axis.text.x = element_text(size = 12, colour = "black"),
           axis.text.y = element_text(size = 12, colour = "black"),
           axis.title.x = element_text(size = 12, colour = "black"),
           plot.title = element_text(size = 12, colour = "black"))

    ggsave(
      filename = paste0("Studies_scripts/figures/bubble_", clean_filename(model_name), "_", mod_term, ".pdf"),
      plot = p,
      width = 6, height = 5, units = "in"
    )
  }
}
```


6. Miscellaneous
```{r}
####A look at the relationship between biome and some of these predictors I used
just_biome <- cleaned_datasets$family.0001

kruskal.test(just_biome$log_fava ~ just_biome$true_biome)
boxplot(log_fava ~ true_biome, data = just_biome,
        xlab = "Biome", ylab = "FAVA (log)",
        main = "FAVA across Biomes", col = "lightgray")

kruskal.test(just_biome$Richprop ~ just_biome$true_biome)
boxplot(Richprop ~ true_biome, data = just_biome,
        xlab = "Biome", ylab = "Proportion of variation due to richness",
        main = "Richness proportion across Biomes", col = "lightgray")

kruskal.test(just_biome$average_turnover ~ just_biome$true_biome)
boxplot(average_turnover ~ true_biome, data = just_biome,
        xlab = "Biome", ylab = "Average turnover",
        main = "Average turnover across biomes", col = "lightgray")

#####DISTRIBUTION OF ARTIFACTS ACROSS BIOMES!!!!!
###Adding these in case I add more analyses later
all_removed
all_removed_notax <- all_removed[,c(11,12)]

reason_counts <- all_removed_notax %>%
  group_by(source_table, reason) %>%
  summarise(count = n(), .groups = "drop") %>%
  pivot_wider(names_from = reason, values_from = count, values_fill = 0)

all_removed_usable_comb <- right_join(
  reason_counts,
  family.0001_complete,
  by = c("source_table" = "ID")
)

reason_cols <- setdiff(names(reason_counts), "source_table")

all_removed_usable_comb[reason_cols] <- lapply(
  all_removed_usable_comb[reason_cols],
  function(x) ifelse(is.na(x), 0, x)
)

all_removed_usable_comb$dynamic <- all_removed_usable_comb$richness - (all_removed_usable_comb$`0_to_1` + all_removed_usable_comb$all_1s + all_removed_usable_comb$`1_to_0`)

plot_data <- all_removed_usable_comb %>%
  select(true_biome, all_1s, `0_to_1`, `1_to_0`, dynamic) %>%
  pivot_longer(cols = c(all_1s, `0_to_1`, `1_to_0`, dynamic),
               names_to = "reason",
               values_to = "count")
plot_data_prop <- plot_data %>%
  group_by(true_biome) %>%
  mutate(prop = count / sum(count)) %>%
  ungroup()

ggplot(plot_data_prop, aes(x = true_biome, y = prop, fill = reason)) +
  geom_bar(stat = "identity") +
  labs(x = "Biome", y = "Proportion", fill = "Reason",
       title = "Proportional Bar Plot of Transition Reasons by Biome") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

contingency_table <- all_removed_usable_comb %>%
  select(true_biome, all_1s, `0_to_1`, `1_to_0`, dynamic) %>%
  pivot_longer(cols = c(all_1s, `0_to_1`, `1_to_0`, dynamic),
               names_to = "reason", values_to = "count") %>%
  group_by(true_biome, reason) %>%
  summarise(count = sum(count), .groups = "drop") %>%
  pivot_wider(names_from = true_biome, values_from = count, values_fill = 0)

# Convert to matrix and run chi-squared test
chisq_result <- chisq.test(as.matrix(contingency_table[,-1]))
chisq_result

######Families
genus_ce_no_art #Issue with the genus data - I forgot to screen for numbered genera so this might create some artifacts
family_ce_no_art
name_updates <- data.frame(old_names = c("Lawsonella", "Intestinimonas", "Howardella"),
                           new_names = c("Lawsonellaceae", "Eubacteriales_Incertae_Sedis(Intestinimonas)", "Clostridiales_Incertae_Sedis"))

names_to_remove <- c("Burkholderiales_noname", "Clostridiales_noname", "Bacteroidales_noname", "Bacteroidales", "Clostridiales", "Candidatus", "butyrate-producing", "Paucibacter", "Gemella", "Epulopiscium")

family_ce_no_art <- family_ce_no_art %>%
  mutate(Group = ifelse(Group %in% name_updates$old_names,
                        name_updates$new_names[match(Group, name_updates$old_names)],
                        Group))

family_ce_no_art <- family_ce_no_art %>%
  filter(!Group %in% names_to_remove)
##Done

#' Compute standardized c- and persistence ranks for each family across subjects
#' and generate a plot showing their median ranks and variation
make_colonization_persistence_plot <- function(df) {
  library(dplyr)
  library(ggplot2)
  library(tidyr)
  library(forcats)

  # Add persistence
  df <- df %>% mutate(persistence = 1 / e)

  # Rank within each subject
  ranked_df <- df %>%
    group_by(ID) %>%
    mutate(
      c_rank = rank(-c, ties.method = "average"),
      p_rank = rank(-persistence, ties.method = "average")
    ) %>%
    group_by(ID) %>%
    mutate(
      c_rank_scaled = (c_rank - min(c_rank)) / (max(c_rank) - min(c_rank)),
      p_rank_scaled = (p_rank - min(p_rank)) / (max(p_rank) - min(p_rank))
    ) %>%
    ungroup()

  # Reshape to long format for plotting
  long_ranked <- ranked_df %>%
    select(Group, ID, c_rank_scaled, p_rank_scaled) %>%
    pivot_longer(cols = c(c_rank_scaled, p_rank_scaled),
                 names_to = "rank_type", values_to = "rank")

  # Study count per family
  family_study_counts <- long_ranked %>%
    group_by(Group) %>%
    summarise(study_count = n_distinct(ID), .groups = "drop")

  # Wilcoxon signed-rank test for c vs. p ranks
  pvals_df <- long_ranked %>%
    pivot_wider(names_from = rank_type, values_from = rank) %>%
    group_by(Group) %>%
    summarise(
      pval = tryCatch(
        wilcox.test(c_rank_scaled, p_rank_scaled, paired = TRUE)$p.value,
        error = function(e) NA_real_
      ), .groups = "drop"
    ) %>%
    mutate(star = case_when(
      is.na(pval) ~ "",
      pval < 0.001 ~ "***",
      pval < 0.01 ~ "**",
      pval < 0.05 ~ "*",
      TRUE ~ ""
    ))

  # Summary stats
  summary_df <- long_ranked %>%
    group_by(Group, rank_type) %>%
    summarise(
      median_rank = median(rank, na.rm = TRUE),
      q25 = quantile(rank, 0.25, na.rm = TRUE),
      q75 = quantile(rank, 0.75, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    left_join(family_study_counts, by = "Group") %>%
    left_join(pvals_df, by = "Group") %>%
    mutate(label = paste0(Group, star, " (n=", study_count, ")"))

  # Order families by c_rank median for plotting
  order_levels <- summary_df %>%
    filter(rank_type == "c_rank_scaled") %>%
    arrange(desc(median_rank)) %>%
    pull(label)
  summary_df$label <- factor(summary_df$label, levels = order_levels)

  # Plot
  p <- ggplot(summary_df, aes(x = median_rank, y = label, color = rank_type)) +
    geom_point(position = position_dodge(width = 0.5)) +
    geom_errorbarh(aes(xmin = q25, xmax = q75), height = 0.3,
                   position = position_dodge(width = 0.5)) +
    scale_color_manual(values = c("c_rank_scaled" = "steelblue", "p_rank_scaled" = "darkorange"),
                       labels = c("Colonization rank", "Persistence rank")) +
    labs(x = "Standardized rank (0 = lowest, 1 = highest)", y = "Family (Study Count)",
         color = "Rank type",
         title = "Colonization and Persistence Ranks by Family") +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 8),
          axis.text.x = element_text(size = 10),
          legend.position = "top")

  return(p)
}


##by biome now
#' Compute standardized c- and persistence ranks for each family across subjects,
#' grouped by biome, and generate a plot per biome
make_colonization_persistence_plots_by_biome <- function(df, biome_lookup) {
  library(dplyr)
  library(ggplot2)
  library(tidyr)
  library(forcats)
  library(purrr)

  # Merge in biome information
  df <- df %>% left_join(biome_lookup, by = "ID")

  # Add persistence
  df <- df %>% mutate(persistence = 1 / e)

  # Rank within each subject
  ranked_df <- df %>%
    group_by(ID) %>%
    mutate(
      c_rank = rank(-c, ties.method = "average"),
      p_rank = rank(-persistence, ties.method = "average")
    ) %>%
    group_by(ID) %>%
    mutate(
      c_rank_scaled = (c_rank - min(c_rank)) / (max(c_rank) - min(c_rank)),
      p_rank_scaled = (p_rank - min(p_rank)) / (max(p_rank) - min(p_rank))
    ) %>%
    ungroup()

  # Reshape to long format
  long_ranked <- ranked_df %>%
    select(true_biome, Group, ID, c_rank_scaled, p_rank_scaled) %>%
    pivot_longer(cols = c(c_rank_scaled, p_rank_scaled),
                 names_to = "rank_type", values_to = "rank")

  # Perform Wilcoxon tests per Group x biome
  pvals_df <- long_ranked %>%
    pivot_wider(names_from = rank_type, values_from = rank) %>%
    group_by(true_biome, Group) %>%
    summarise(
      pval = tryCatch(
        wilcox.test(c_rank_scaled, p_rank_scaled, paired = TRUE)$p.value,
        error = function(e) NA_real_
      ), .groups = "drop"
    ) %>%
    mutate(star = case_when(
      is.na(pval) ~ "",
      pval < 0.001 ~ "***",
      pval < 0.01 ~ "**",
      pval < 0.05 ~ "*",
      TRUE ~ ""
    ))

  # Study count per family per biome
  family_study_counts <- long_ranked %>%
    group_by(true_biome, Group) %>%
    summarise(study_count = n_distinct(ID), .groups = "drop")

  # Summary stats
  summary_df <- long_ranked %>%
    group_by(true_biome, Group, rank_type) %>%
    summarise(
      median_rank = median(rank, na.rm = TRUE),
      q25 = quantile(rank, 0.25, na.rm = TRUE),
      q75 = quantile(rank, 0.75, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    left_join(family_study_counts, by = c("true_biome", "Group")) %>%
    left_join(pvals_df, by = c("true_biome", "Group")) %>%
    mutate(label = paste0(Group, star, " (n=", study_count, ")"))

  # Split by biome and return a list of ggplot objects
  plots_by_biome <- summary_df %>%
    split(.$true_biome) %>%
    map(function(biome_df) {
      # order labels by c_rank median
      order_levels <- biome_df %>%
        filter(rank_type == "c_rank_scaled") %>%
        arrange(desc(median_rank)) %>%
        pull(label)
      biome_df$label <- factor(biome_df$label, levels = order_levels)

      ggplot(biome_df, aes(x = median_rank, y = label, color = rank_type)) +
        geom_point(position = position_dodge(width = 0.5)) +
        geom_errorbarh(aes(xmin = q25, xmax = q75), height = 0.3,
                       position = position_dodge(width = 0.5)) +
        scale_color_manual(values = c("c_rank_scaled" = "steelblue", "p_rank_scaled" = "darkorange"),
                           labels = c("Colonization rank", "Persistence rank")) +
        labs(x = "Standardized rank (0 = lowest, 1 = highest)", y = "Family (Study Count)",
             color = "Rank type",
             title = paste("Colonization and Persistence Ranks:", unique(biome_df$true_biome))) +
        theme_minimal() +
        theme(
          axis.text.y = element_text(size = 7),
          axis.text.x = element_text(size = 10),
          legend.position = "top"
        )
    })

  return(plots_by_biome)
}

#' Plot colonization and persistence ranks across all subjects
#' Only includes families with significant paired differences
make_colonization_persistence_plot_sigonly <- function(df) {
  library(dplyr)
  library(ggplot2)
  library(tidyr)
  library(forcats)

  # Add persistence
  df <- df %>% mutate(persistence = 1 / e)

  # Rank within each subject
  ranked_df <- df %>%
    group_by(ID) %>%
    mutate(
      c_rank = rank(-c, ties.method = "average"),
      p_rank = rank(-persistence, ties.method = "average")
    ) %>%
    group_by(ID) %>%
    mutate(
      c_rank_scaled = (c_rank - min(c_rank)) / (max(c_rank) - min(c_rank)),
      p_rank_scaled = (p_rank - min(p_rank)) / (max(p_rank) - min(p_rank))
    ) %>%
    ungroup()

  # Reshape to long format for plotting
  long_ranked <- ranked_df %>%
    select(Group, ID, c_rank_scaled, p_rank_scaled) %>%
    pivot_longer(cols = c(c_rank_scaled, p_rank_scaled),
                 names_to = "rank_type", values_to = "rank")

  # Study count per family
  family_study_counts <- long_ranked %>%
    group_by(Group) %>%
    summarise(study_count = n_distinct(ID), .groups = "drop")

  # Wilcoxon signed-rank test for c vs. p ranks
  pvals_df <- long_ranked %>%
    pivot_wider(names_from = rank_type, values_from = rank) %>%
    group_by(Group) %>%
    summarise(
      pval = tryCatch(
        wilcox.test(c_rank_scaled, p_rank_scaled, paired = TRUE)$p.value,
        error = function(e) NA_real_
      ), .groups = "drop"
    ) %>%
    mutate(star = case_when(
      is.na(pval) ~ "",
      pval < 0.001 ~ "***",
      pval < 0.01 ~ "**",
      pval < 0.05 ~ "*",
      TRUE ~ ""
    )) %>%
    filter(star != "")  # keep only significant

  # Summary stats
  summary_df <- long_ranked %>%
    group_by(Group, rank_type) %>%
    summarise(
      median_rank = median(rank, na.rm = TRUE),
      q25 = quantile(rank, 0.25, na.rm = TRUE),
      q75 = quantile(rank, 0.75, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    left_join(family_study_counts, by = "Group") %>%
    left_join(pvals_df, by = "Group") %>%
    filter(!is.na(star)) %>%
    mutate(label = paste0(Group, star, " (n=", study_count, ")"))

  # Order families by c_rank median for plotting
  order_levels <- summary_df %>%
    filter(rank_type == "c_rank_scaled") %>%
    arrange(desc(median_rank)) %>%
    pull(label)
  summary_df$label <- factor(summary_df$label, levels = order_levels)

  # Plot
  p <- ggplot(summary_df, aes(x = median_rank, y = label, color = rank_type)) +
    geom_point(position = position_dodge(width = 0.5)) +
    geom_errorbarh(aes(xmin = q25, xmax = q75), height = 0.3,
                   position = position_dodge(width = 0.5)) +
    scale_color_manual(values = c("c_rank_scaled" = "steelblue", "p_rank_scaled" = "darkorange"),
                       labels = c("Colonization", "Persistence")) +
    labs(x = "Standardized rank", y = "Family (Study Count)",
         color = "Rank type",
         title = "Colonization/persistence ranks"
         ) +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 8),
          axis.text.x = element_text(size = 10),
          legend.position = "top")

  return(p)
}


#' Plot per-biome colonization/persistence ranks for significantly different families
make_colonization_persistence_plots_by_biome_sigonly <- function(df, biome_lookup) {
  library(dplyr)
  library(ggplot2)
  library(tidyr)
  library(forcats)
  library(purrr)

  # Merge in biome information
  df <- df %>% left_join(biome_lookup, by = "ID")
  df <- df %>% mutate(persistence = 1 / e)

  ranked_df <- df %>%
    group_by(ID) %>%
    mutate(
      c_rank = rank(-c, ties.method = "average"),
      p_rank = rank(-persistence, ties.method = "average")
    ) %>%
    group_by(ID) %>%
    mutate(
      c_rank_scaled = (c_rank - min(c_rank)) / (max(c_rank) - min(c_rank)),
      p_rank_scaled = (p_rank - min(p_rank)) / (max(p_rank) - min(p_rank))
    ) %>%
    ungroup()

  long_ranked <- ranked_df %>%
    select(true_biome, Group, ID, c_rank_scaled, p_rank_scaled) %>%
    pivot_longer(cols = c(c_rank_scaled, p_rank_scaled),
                 names_to = "rank_type", values_to = "rank")

  # Wilcoxon tests per group per biome
  pvals_df <- long_ranked %>%
    pivot_wider(names_from = rank_type, values_from = rank) %>%
    group_by(true_biome, Group) %>%
    summarise(pval = tryCatch(
      wilcox.test(c_rank_scaled, p_rank_scaled, paired = TRUE)$p.value,
      error = function(e) NA_real_
    ), .groups = "drop") %>%
    mutate(star = case_when(
      is.na(pval) ~ "",
      pval < 0.001 ~ "***",
      pval < 0.01 ~ "**",
      pval < 0.05 ~ "*",
      TRUE ~ ""
    )) %>%
    filter(star != "")

  family_study_counts <- long_ranked %>%
    group_by(true_biome, Group) %>%
    summarise(study_count = n_distinct(ID), .groups = "drop")

  summary_df <- long_ranked %>%
    group_by(true_biome, Group, rank_type) %>%
    summarise(
      median_rank = median(rank, na.rm = TRUE),
      q25 = quantile(rank, 0.25, na.rm = TRUE),
      q75 = quantile(rank, 0.75, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    left_join(family_study_counts, by = c("true_biome", "Group")) %>%
    left_join(pvals_df, by = c("true_biome", "Group")) %>%
    filter(!is.na(star)) %>%
    mutate(label = paste0(Group, star, " (n=", study_count, ")"))

  plots_by_biome <- summary_df %>%
    split(.$true_biome) %>%
    map(function(biome_df) {
      order_levels <- biome_df %>%
        filter(rank_type == "c_rank_scaled") %>%
        arrange(desc(median_rank)) %>%
        pull(label)
      biome_df$label <- factor(biome_df$label, levels = order_levels)

      ggplot(biome_df, aes(x = median_rank, y = label, color = rank_type)) +
        geom_point(position = position_dodge(width = 0.5)) +
        geom_errorbarh(aes(xmin = q25, xmax = q75), height = 0.3,
                       position = position_dodge(width = 0.5)) +
        scale_color_manual(values = c("c_rank_scaled" = "steelblue", "p_rank_scaled" = "darkorange"),
                           labels = c("Colonization", "Persistence")) +
        labs(x = "Standardized rank", y = "Family (Study Count)",
             color = "Rank type",
             title = paste("Colonization/persistence ranks", unique(biome_df$true_biome))) +
        theme_minimal() +
        theme(
          axis.text.y = element_text(size = 7),
          axis.text.x = element_text(size = 10),
          legend.position = "top"
        )
    })

  return(plots_by_biome)
}


make_colonization_persistence_plot(family_ce_no_art)
make_colonization_persistence_plots_by_biome(family_ce_no_art, family.0001_complete)
make_colonization_persistence_plot_sigonly(family_ce_no_art)
make_colonization_persistence_plots_by_biome_sigonly(family_ce_no_art, family.0001_complete)

##########Summarizing reasons for removal by family
#' Create proportional bar plot showing reasons for removal per family
plot_removal_reason_by_family <- function(df, df_biome) {
  library(dplyr)
  library(ggplot2)
  library(tidyr)

  df <- df %>% left_join(df_biome, by = c("source_table" = "ID")) %>% filter(!is.na(true_biome))

  df_clean <- df %>%
    filter(!is.na(family) & family != "") %>%
    group_by(family, reason) %>%
    summarise(n = n(), .groups = "drop") %>%
    group_by(family) %>%
    mutate(
      total = sum(n),
      prop = n / total,
      label = paste0(family, " (n=", total, ")")
    ) %>%
    ungroup()

  ggplot(df_clean, aes(x = prop, y = reorder(label, prop), fill = reason)) +
    geom_col(position = "stack") +
    labs(x = "Proportion of removals", y = "Family (count)", fill = "Reason",
         title = "Removal Reason Breakdown by Family") +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 8))
}


#' Create proportional bar plot faceted by biome (requires joining)
plot_removal_reason_by_family_biome <- function(df_removal, df_biome) {
  library(dplyr)
  library(ggplot2)
  library(tidyr)

  df_joined <- df_removal %>%
    left_join(df_biome, by = c("source_table" = "ID")) %>%
    filter(!is.na(true_biome) & !is.na(family) & family != "")

  df_summary <- df_joined %>%
    group_by(true_biome, family, reason) %>%
    summarise(n = n(), .groups = "drop") %>%
    group_by(true_biome, family) %>%
    mutate(
      total = sum(n),
      prop = n / total,
      label = paste0(family, " (n=", total, ")")
    ) %>%
    ungroup()

  ggplot(df_summary, aes(x = prop, y = reorder(label, prop), fill = reason)) +
    geom_col(position = "stack") +
    facet_wrap(~true_biome, scales = "free_y") +
    labs(x = "Proportion of removals", y = "Family (count)", fill = "Reason",
         title = "Removal Reasons by Family and Biome") +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 6))
}

##removing some families that aren't families
name_updates <- data.frame(old_names = c("Lawsonella", "Intestinimonas", "Howardella"),
                           new_names = c("Lawsonellaceae", "Eubacteriales_Incertae_Sedis(Intestinimonas)", "Clostridiales_Incertae_Sedis"))

names_to_remove <- c("Burkholderiales_noname", "Clostridiales_noname", "Bacteroidales_noname", "Bacteroidales", "Clostridiales", "Candidatus", "butyrate-producing", "Paucibacter", "Gemella", "Epulopiscium")

all_removed <- all_removed %>%
  mutate(family = ifelse(family %in% name_updates$old_names,
                        name_updates$new_names[match(family, name_updates$old_names)],
                        family))

all_removed <- all_removed %>%
  filter(!family %in% names_to_remove)

plot_removal_reason_by_family(all_removed, family.0001_complete)
plot_removal_reason_by_family_biome(all_removed, family.0001_complete)



```

```{r}
###goofing around
view(family.0001_complete)

standardize_columns <- function(df) {
  # Rename columns if present
  if ("distinct_families" %in% names(df)) {
    names(df)[names(df) == "distinct_families"] <- "distinct_taxa"
  }
  if ("distinct_genera" %in% names(df)) {
    names(df)[names(df) == "distinct_genera"] <- "distinct_taxa"
  }
  return(df)
}

# Apply to all datasets
datasets_standardized <- lapply(datasets, standardize_columns)

standard_cols <- colnames(datasets_standardized$family.0001[,5:16])

test <- lapply(datasets_standardized, function(df) {
  # Create a new empty row with standard columns filled with NAs
  out <- setNames(as.data.frame(matrix(NA, nrow = 1, ncol = length(standard_cols))), standard_cols)
  
  # If the target row exists:
  row <- df[df$ID == "M2184_F4_skin.extra", ] #replicate with M3659_Person.03.3_Foot
  if (nrow(row) > 0) {
    # Fill in the matching columns
    matching_cols <- intersect(names(row), standard_cols)
    out[1, matching_cols] <- row[1, matching_cols]
    return(out)
  } else {
    return(NULL)
  }
})

test_rows <- do.call(rbind, test)

#Do with both M2184_M3_skin.extra and M6638_X1398.1.1.BMIF_nasopharyngeal
palm_fam.0001 <- family_ce_info[which(family_ce_info$ID == "M2184_M3_skin.extra"),]
palm_fam_noart <-  family_ce_no_art[which(family_ce_no_art$ID == "M2184_M3_skin.extra"),]
palm_gen.0001 <- genus_ce_info[which(genus_ce_info$ID == "M2184_M3_skin.extra"),]
palm_gen_noart <- genus_ce_no_art[which(genus_ce_no_art$ID == "M2184_M3_skin.extra"),]

palm_fam.0001$label <- "M3 palm (family)"
palm_fam_noart$label <- "M3 palm (family, no artifacts)"
palm_gen.0001$label <- "M3 palm (genus)"
palm_gen_noart$label <- "M3 palm (genus, no artifacts)"

# Combine them
combined_plot_data <- bind_rows(
  palm_fam.0001,
  palm_fam_noart,
  palm_gen.0001,
  palm_gen_noart
)

combined_plot_data_fam <- bind_rows(
  palm_fam.0001,
  palm_fam_noart
)

# Create the plot
ggplot(combined_plot_data, aes(x = log(c), y = log(1/e))) +
  geom_point(alpha = 0.7) +
  facet_wrap(~ label, scales = "free") +
  labs(
    x = "Colonization (log)",
    y = "Persistence (log)",
    title = NULL
  ) +
  theme_minimal() +
  theme(strip.text = element_text(size = 10, face = "bold"))

ggplot(combined_plot_data_fam, aes(x = log(c), y = log(1/e))) +
  geom_point(alpha = 0.7) +
  facet_wrap(~ label, scales = "free") +
  labs(
    x = "Colonization (log)",
    y = "Persistence (log)",
    title = NULL
  ) +
  theme_minimal() +
  theme(strip.text = element_text(size = 10, face = "bold"))


#Now with M6638_X1398.1.1.BMIF_nasopharyngeal
nasopharynx_fam.0001 <- family_ce_info[which(family_ce_info$ID == "M6638_X1398.1.1.BMIF_nasopharyngeal"),]
nasopharynx_fam_noart <-  family_ce_no_art[which(family_ce_no_art$ID == "M6638_X1398.1.1.BMIF_nasopharyngeal"),]
nasopharynx_gen.0001 <- genus_ce_info[which(genus_ce_info$ID == "M6638_X1398.1.1.BMIF_nasopharyngeal"),]
nasopharynx_gen_noart <- genus_ce_no_art[which(genus_ce_no_art$ID == "M6638_X1398.1.1.BMIF_nasopharyngeal"),]

nasopharynx_fam.0001$label <- "1398 nasopharynx (family)"
nasopharynx_fam_noart$label <- "1398 nasopharynx (family, no artifacts)"
nasopharynx_gen.0001$label <- "1398 nasopharynx (genus)"
nasopharynx_gen_noart$label <- "1398 nasopharynx (genus, no artifacts)"

# Combine them
combined_plot_data <- bind_rows(
  nasopharynx_fam.0001,
  nasopharynx_fam_noart,
  nasopharynx_gen.0001,
  nasopharynx_gen_noart
)

combined_plot_data_fam <- bind_rows(
  nasopharynx_fam.0001,
  nasopharynx_fam_noart
)

# Create the plot
ggplot(combined_plot_data, aes(x = log(c), y = log(1/e))) +
  geom_point(alpha = 0.7) +
  facet_wrap(~ label, scales = "free") +
  labs(
    x = "Colonization (log)",
    y = "Persistence (log)",
    title = NULL
  ) +
  theme_minimal() +
  theme(strip.text = element_text(size = 10, face = "bold"))

ggplot(combined_plot_data_fam, aes(x = log(c), y = log(1/e))) +
  geom_point(alpha = 0.7) +
  facet_wrap(~ label, scales = "free") +
  labs(
    x = "Colonization (log)",
    y = "Persistence (log)",
    title = NULL
  ) +
  theme_minimal() +
  theme(strip.text = element_text(size = 10, face = "bold"))

#M5196_UT39_vaginal
vagina_fam.0001 <- family_ce_info[which(family_ce_info$ID == "M5196_UT39_vaginal"),]
vagina_fam_noart <-  family_ce_no_art[which(family_ce_no_art$ID == "M5196_UT39_vaginal"),]
vagina_gen.0001 <- genus_ce_info[which(genus_ce_info$ID == "M5196_UT39_vaginal"),]
vagina_gen_noart <- genus_ce_no_art[which(genus_ce_no_art$ID == "M5196_UT39_vaginal"),]

vagina_fam.0001$label <- "UT39 vagina (family)"
vagina_fam_noart$label <- "UT39 vagina (family, no artifacts)"
vagina_gen.0001$label <- "UT39 vagina (genus)"
vagina_gen_noart$label <- "UT39 vagina (genus, no artifacts)"

# Combine them
combined_plot_data <- bind_rows(
  vagina_fam.0001,
  vagina_fam_noart,
  vagina_gen.0001,
  vagina_gen_noart
)

combined_plot_data_fam <- bind_rows(
  vagina_fam.0001,
  vagina_fam_noart
)

# Create the plot
ggplot(combined_plot_data, aes(x = log(c), y = log(1/e))) +
  geom_point(alpha = 0.7) +
  facet_wrap(~ label, scales = "free") +
  labs(
    x = "Colonization (log)",
    y = "Persistence (log)",
    title = NULL
  ) +
  theme_minimal() +
  theme(strip.text = element_text(size = 10, face = "bold"))

ggplot(combined_plot_data_fam, aes(x = log(c), y = log(1/e))) +
  geom_point(alpha = 0.7) +
  facet_wrap(~ label, scales = "free") +
  labs(
    x = "Colonization (log)",
    y = "Persistence (log)",
    title = NULL
  ) +
  theme_minimal() +
  theme(strip.text = element_text(size = 10, face = "bold"))

```

```{r}

#dealing with NAs in my binary moderators (replacing with 0s for now)
cleaned_datasets <- lapply(datasets_slim, function(df) {
  # Replace NAs with 0s in binary columns
  bin_vars <- c("disease", "sustained_antibiotic", "sustained_probiotic", "development", "pregnancy")
  df[bin_vars] <- lapply(df[bin_vars], function(x) {
    x[is.na(x)] <- 0
    return(x)
  })

  # Combine them into a string based on which are 1
  df$disturbance <- apply(df[bin_vars], 1, function(row) {
    active_flags <- names(row)[which(row == 1)]
    if (length(active_flags) == 0) {
      return("none")  # fallback if no flags are on
    } else {
      return(paste(active_flags, collapse = "_"))
    }
  })

  # Ensure study_ID is a factor
  df$study_ID <- as.factor(df$study_ID)
  df$log_timepoint <- log(df$num_timepoint)
  df$inverse_timepoint <- 1/(df$log_timepoint)
  df$log_elapsed <- log(df$time_elapsed)
  df$log_turnover <- log(df$average_turnover)
  df$log_fava <- log(df$fava)
  df$time_standard <- df$time_elapsed/df$num_timepoint
  df$sqrt_freq <- sqrt(df$time_standard)
  df$sqrt_ste <- sqrt(df$ste)
  
  df$true_biome <- relevel(
  factor(df$true_biome),
  ref = "gut"
)

  return(df)
})


###Doing some multivariate models:
model <- rma.mv(
  yi = slope,       
  V = ste,        
  mods = ~ true_biome*disease + sustained_antibiotic + sustained_probiotic + development + pregnancy + log_fava + pub_year_simple + sqrt_ste + sqrt_freq,  
  random = list(~1 | study_ID/subject_id),   
  data = cleaned_datasets$family.no_art,
  method = "REML",  
  test = "t",
  sparse = TRUE
)

model <- rma.mv(
  yi = slope,       
  V = ste,        
  mods = ~ pub_year_simple + sqrt_ste + sqrt_freq,  
  random = list(~1 | study_ID/subject_id),   
  data = cleaned_datasets$family.no_art,
  method = "REML",  
  test = "t",
  sparse = TRUE
)

summary(model)
summary(lm(fava ~ log(num_timepoint), data = cleaned_datasets$family.no_art))
hist(cleaned_datasets$family.no_art$fava)
bubble_plot(model,
                     mod = "average_turnover",
                     xlab = "average_turnover",
                     ylab = "slope",
                     group = "study_ID",
                     k = TRUE, g = TRUE,
                     legend.pos = "top.left") +
         ggtitle(paste("Bubble_plot")) +
         theme(
           axis.text.x = element_text(size = 12, colour = "black"),
           axis.text.y = element_text(size = 12, colour = "black"),
           axis.title.x = element_text(size = 12, colour = "black"),
           plot.title = element_text(size = 12, colour = "black"))

#running it on all of them:
run_meta_model <- function(data, effect_size_col, variance_col, level, no_art = FALSE) {
  # Filter data based on minimum number of taxa
  min_taxa_col <- ifelse(level == "family", "distinct_families", "distinct_genera")
  data_filtered <- data[data[[min_taxa_col]] > 9, ]
  
  # Model label
  art_status <- ifelse(no_art, "no_art", "0001")
  model_label <- paste0("mod_", ifelse(effect_size_col == "z", "ML_Z", "slope"),
                        "_full_", substr(level, 1, 3), ".", art_status)
  
  cat("##########################\n")
  cat(model_label, "\n")

  # Try sparse model first
  try_sparse <- try({
    model <- rma.mv(
      yi = data_filtered[[effect_size_col]],
      V = data_filtered[[variance_col]],
      mods = ~ true_biome*disease + sustained_antibiotic + sustained_probiotic + development + pregnancy + log_fava,
      random = list(~1 | study_ID/subject_id),
      method = "REML",
      data = data_filtered,
      sparse = TRUE,
      test = "t"
    )
  }, silent = TRUE)

  # Retry without sparse if error
  if (inherits(try_sparse, "try-error")) {
    cat("Model failed with sparse = TRUE, retrying with sparse = FALSE...\n")
    model <- rma.mv(
      yi = data_filtered[[effect_size_col]],
      V = data_filtered[[variance_col]],
      mods = ~ true_biome*disease + sustained_antibiotic + sustained_probiotic + development + pregnancy + log_fava,
      random = list(~1 | study_ID/subject_id),
      method = "REML",
      data = data_filtered,
      sparse = FALSE,
      test = "t"
    )
  } else {
    model <- try_sparse
  }

  print(summary(model))
  return(model)
}

full_models <- list()

for (name in names(cleaned_datasets)) {
  dat <- cleaned_datasets[[name]]
  level <- ifelse(grepl("genus", name), "genus", "family")
  no_art <- grepl("no_art", name)
  
  # Z effect size (e.g., transformed correlation)
  full_models[[paste0("ML_Z_", name)]] <- run_meta_model(
    dat, effect_size_col = "z", variance_col = "v", level = level, no_art = no_art)
  
  # Slope effect size
  full_models[[paste0("slope_", name)]] <- run_meta_model(
    dat, effect_size_col = "slope", variance_col = "ste", level = level, no_art = no_art)
}

full_meta_models <- extract_model_summary(full_models)

write.csv(full_meta_models, "multiple_metaregression_outputs.csv")

```

Adding facet wrapped figures for FAVA and freq, pub_year
```{r}
ext_models <- list()

moderators <- list(
  log_fava = "FAVA (log)",
  sqrt_freq = "Sampling Frequency (sqrt)",
  pub_year_simple = "Publication year"
)

for (name in names(cleaned_datasets)) {
  dat <- cleaned_datasets[[name]]
  level <- ifelse(grepl("genus", name), "genus", "family")

  for (mod_name in names(moderators)) {
    # Rho (z)
    model_z <- run_meta_model_diversity(dat, moderator = mod_name, level = level, effect_size_col = "z", variance_col = "v")
    ext_models[[paste0("diversity_", mod_name, "_z_", name)]] <- model_z

    # Slope (beta)
    model_slope <- run_meta_model_diversity(dat, moderator = mod_name, level = level, effect_size_col = "slope", variance_col = "ste")
    ext_models[[paste0("diversity_", mod_name, "_slope_", name)]] <- model_slope
  }
}

# Save summaries if needed
ext_summary <- extract_model_summary(ext_models)
all_bubbles <- c(ext_models, small_models)

facet_bubble_plots_by_moderator(ext_models)

cleaned_datasets$family.0001$pub_year

```

Bias on pipeline type and sequencing platform
```{r}
pipeline_platform <- read.csv("Final_study_list.csv")
#view(pipeline_platform)
pipeline_platform <- pipeline_platform[,c(3,4,8,9)]
cleaned_datasets_pipelines <- lapply(cleaned_datasets, function(df){
  df_merged <- df %>%
    left_join(pipeline_platform, by = c("study_ID" = "study_code"))
  df_merged$diseases <- df_merged$disease
  df_merged$diseases[which(df_merged$diseases == 0)] <- "Healthy"
  df_merged$diseases[which(df_merged$diseases == 1)] <- "Disease"
  df_merged$diseases <- relevel(
  factor(df_merged$diseases),
  ref = "Healthy")
  return(df_merged)
})

model <- rma.mv(
  yi = z,       
  V = v,        
  mods = ~ Analysis_pipeline + Sequencing_platform + experiment_type,  
  random = list(~1 | study_ID/subject_id),   
  data = cleaned_datasets_pipelines$family.no_art,
  method = "REML",  
  test = "t",
  sparse = TRUE
)

#BIAS TESTING FINAL
bias_models <- list()
moderators <- list(
  Analysis_pipeline = "pipeline",
  Sequencing_platform = "platform",
  sqrt_freq = "Sampling Frequency (sqrt)",
  pub_year_simple = "Publication year",
  experiment_type = "experiment type"
)

for (name in names(cleaned_datasets_pipelines)) {
  dat <- cleaned_datasets_pipelines[[name]]
  level <- ifelse(grepl("genus", name), "genus", "family")

  for (mod_name in names(moderators)) {
    # Rho (z)
    model_z <- run_meta_model_diversity(dat, moderator = mod_name, level = level, effect_size_col = "z", variance_col = "v")
    bias_models[[paste0("diversity_", mod_name, "_z_", name)]] <- model_z

    # Slope (beta)
    model_slope <- run_meta_model_diversity(dat, moderator = mod_name, level = level, effect_size_col = "slope", variance_col = "ste")
    bias_models[[paste0("diversity_", mod_name, "_slope_", name)]] <- model_slope
  }
}

# Save summaries if needed
all_biases <- c(bias_models, small_models)
bias_models_summary <- extract_model_summary(bias_models)
all_bias_summary <- rbind(bias_models_summary, small_model_table)
write.csv(all_bias_summary, "all_bias_pipelines.csv")
view(bias_models_summary)

```

Running the model with the biases
```{r}
run_meta_model_bias <- function(data, effect_size_col, variance_col, level, no_art = FALSE) {
  # Filter data based on minimum number of taxa
  min_taxa_col <- ifelse(level == "family", "distinct_families", "distinct_genera")
  data_filtered <- data[data[[min_taxa_col]] > 9, ]
  
  # Model label
  art_status <- ifelse(no_art, "no_art", "0001")
  model_label <- paste0("mod_", ifelse(effect_size_col == "z", "ML_Z", "slope"),
                        "_full_", substr(level, 1, 3), ".", art_status)
  
  cat("##########################\n")
  cat(model_label, "\n")
  data_filtered$sqrt_v <- sqrt(data_filtered[[variance_col]])

  # Try sparse model first
  try_sparse <- try({
    model <- rma.mv(
      yi = data_filtered[[effect_size_col]],
      V = data_filtered[[variance_col]],
      mods = ~ true_biome*disease + sustained_antibiotic + sustained_probiotic + development + pregnancy + log_fava + pub_year_simple + sqrt_freq + experiment_type,
      random = list(~1 | study_ID/subject_id),
      method = "REML",
      data = data_filtered,
      sparse = TRUE,
      test = "t"
    )
  }, silent = TRUE)

  # Retry without sparse if error
  if (inherits(try_sparse, "try-error")) {
    cat("Model failed with sparse = TRUE, retrying with sparse = FALSE...\n")
    model <- rma.mv(
      yi = data_filtered[[effect_size_col]],
      V = data_filtered[[variance_col]],
      mods = ~ true_biome*disease + sustained_antibiotic + sustained_probiotic + development + pregnancy + log_fava + pub_year_simple + sqrt_freq + sqrt_v + experiment_type,
      random = list(~1 | study_ID/subject_id),
      method = "REML",
      data = data_filtered,
      sparse = FALSE,
      test = "t"
    )
  } else {
    model <- try_sparse
  }

  print(summary(model))
  return(model)
}

full_models_bias <- list()

for (name in names(cleaned_datasets_pipelines)) {
  dat <- cleaned_datasets_pipelines[[name]]
  level <- ifelse(grepl("genus", name), "genus", "family")
  no_art <- grepl("no_art", name)
  
  # Z effect size (e.g., transformed correlation)
  full_models_bias[[paste0("ML_Z_", name)]] <- run_meta_model_bias(
    dat, effect_size_col = "z", variance_col = "v", level = level, no_art = no_art)
  
  # Slope effect size
  full_models_bias[[paste0("slope_", name)]] <- run_meta_model_bias(
    dat, effect_size_col = "slope", variance_col = "ste", level = level, no_art = no_art)
}

full_meta_models_bias <- extract_model_summary(full_models_bias)

write.csv(full_meta_models_bias, "multiple_metaregression_bias_outputs.csv")


```

Last thing: rerunning the meta-analytic means with the new biases
```{r}
run_meta_model_bias <- function(data, level = "family", effect_size_col = "z", variance_col = "v") {
  # Filter based on minimum taxa
  min_taxa_col <- ifelse(level == "family", "distinct_families", "distinct_genera")
  data_filtered <- data[data[[min_taxa_col]] > 9, ]
  data_filtered$sqrt_v <- sqrt(data_filtered[[variance_col]])

  # Fit multilevel model
  model <- rma.mv(
    yi = data_filtered[[effect_size_col]],
    V = data_filtered[[variance_col]],
    mods = ~ pub_year_simple + sqrt_freq + experiment_type,
    random = list(
      ~1 | study_ID,
      ~1 | subject_id,
      ~1 | biome_ID
    ),
    method = "REML",
    data = data_filtered,
    sparse = TRUE
  )

  return(model)
}

meta_models_bias <- list()

for (name in names(cleaned_datasets_pipelines)) {
  dat <- cleaned_datasets_pipelines[[name]]
  level <- ifelse(grepl("genus", name), "genus", "family")

  # Rho model
  model_rho <- run_meta_model_bias(dat, level = level, effect_size_col = "z", variance_col = "v")
  model_name_rho <- paste0("mod_rho_", name)
  meta_models_bias[[model_name_rho]] <- model_rho

  # Slope model
  model_slope <- run_meta_model_bias(dat, level = level, effect_size_col = "slope", variance_col = "ste")
  model_name_slope <- paste0("mod_slope_", name)
  meta_models_bias[[model_name_slope]] <- model_slope
}

meta_model_bias_summary <- extract_model_summary(meta_models_bias)

write.csv(meta_model_bias_summary, "meta_means_bias.csv")

```

Making some plots to better understand the multiple regression results
```{r}
df <- read.csv("multiple_metaregression_outputs.csv")
model_names <- c(rep("Effect Size Z: Family (Artefacts)", 14), rep("Effect Size Beta: Family (Artefacts)", 14), rep("Effect Size Z: Family", 14), rep("Effect Size Beta: Family", 14), rep("Effect Size Z: Genus (Artefacts)", 14), rep("Effect Size Beta: Genus (Artefacts)", 14), rep("Effect Size Z: Genus", 14), rep("Effect Size Beta: Genus", 14))
df$model_name <- model_names
df <- df[which(str_detect(df$model_name, "Artefacts") == F),]

df <- df %>%
  mutate(
    term = fct_recode(term,
      "Intercept" = "intrcpt",
      "Body Site: Oral" = "true_biomeoral",
      "Body Site: Resp." = "true_biomerespiratory",
      "Body Site: Skin" = "true_biomeskin",
      "Body Site: Vaginal" = "true_biomevaginal",
      "Disease" = "disease",
      "Antibiotics" = "sustained_antibiotic",
      "Probiotics" = "sustained_probiotic",
      "Infancy" = "development",
      "Pregnancy" = "pregnancy",
      "Log(FAVA)" = "log_fava",
      "Oral × Disease" = "true_biomeoral:disease",
      "Resp. × Disease" = "true_biomerespiratory:disease",
      "Vaginal × Disease" = "true_biomevaginal:disease"
    ),
    term = factor(term, levels = rev(unique(term)))  # reverse for top-down
  )

make_zval_heatmap <- function(df) {
  df <- df %>%
    mutate(
      sig = ifelse(pval < 0.05, "*", ""),
      term = factor(term, levels = rev(unique(term)))  # reverse for top-down
    )

  ggplot(df, aes(x = model_name, y = term)) +
    geom_tile(aes(fill = zval), color = "white") +
    geom_text(aes(label = sig), size = 4, vjust = 0.5) +
    scale_fill_gradient2(
      low = "blue", mid = "white", high = "red", 
      midpoint = 0, limits = c(-max(abs(df$zval)), max(abs(df$zval))),
      name = "z-score"
    ) +
    labs(x = NULL, y = NULL) +
    theme_minimal(base_size = 12) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      panel.grid = element_blank()
    )
}

#with biases checked now

make_zval_heatmap(df)

df2 <- read.csv("multiple_metaregression_bias_outputs.csv")
df2$model_name <- c(rep("Effect Size Z: Family (Artefacts)", 18), rep("Effect Size Beta: Family (Artefacts)", 18), rep("Effect Size Z: Family", 18), rep("Effect Size Beta: Family", 18), rep("Effect Size Z: Genus (Artefacts)", 17), rep("Effect Size Beta: Genus (Artefacts)", 17), rep("Effect Size Z: Genus", 17), rep("Effect Size Beta: Genus", 17))

df2 <- df2 %>%
  mutate(
    term = fct_recode(term,
      "Intercept" = "intrcpt",
      "Body Site: Oral" = "true_biomeoral",
      "Body Site: Resp." = "true_biomerespiratory",
      "Body Site: Skin" = "true_biomeskin",
      "Body Site: Vaginal" = "true_biomevaginal",
      "Disease" = "disease",
      "Antibiotics" = "sustained_antibiotic",
      "Probiotics" = "sustained_probiotic",
      "Infancy" = "development",
      "Pregnancy" = "pregnancy",
      "Log(FAVA)" = "log_fava",
      "Publication Year" = "pub_year_simple",
      "Sampling Freq." = "sqrt_freq",
      "Exp. Type: Assembly" = "experiment_typeAssembly",
      "Exp. Type: Metagenomic" = "experiment_typeMetagenomic",
      "Oral × Disease" = "true_biomeoral:disease",
      "Resp. × Disease" = "true_biomerespiratory:disease",
      "Vaginal × Disease" = "true_biomevaginal:disease"
    ),
    term = factor(term, levels = rev(unique(term)))  # reverse for top-down
  )


df2 <- df2[which(str_detect(df2$model_name, "Artefacts") == F),]
make_zval_heatmap(df2)

```

Making figures for disease as a moderator
```{r}
run_meta_model_disease <- function(data, level = "family", effect_size_col = "z", variance_col = "v") {
  # Filter based on minimum taxa
  min_taxa_col <- ifelse(level == "family", "distinct_families", "distinct_genera")
  data_filtered <- data[data[[min_taxa_col]] > 9, ]
  data_filtered$sqrt_v <- sqrt(data_filtered[[variance_col]])
  data_filtered$biome_disease <- interaction(data_filtered$true_biome, data_filtered$diseases, sep = ":")
  # Prettify interaction term
data_filtered$biome_disease_label <- paste0(as.character(data_filtered$true_biome),
  ": ", as.character(data_filtered$diseases)
)

levels_ordered <- c(
  "skin: Healthy",
  "gut: Healthy", "gut: Disease",
  "oral: Healthy", "oral: Disease",
  "respiratory: Healthy", "respiratory: Disease",
  "vaginal: Healthy", "vaginal: Disease"
)

data_filtered$biome_disease_label <- factor(
  data_filtered$biome_disease_label,
  levels = levels_ordered
)

  # Fit multilevel model
  model <- rma.mv(
    yi = data_filtered[[effect_size_col]],
    V = data_filtered[[variance_col]],
    mods = ~ biome_disease_label -1, ##messing around
    random = list(
      ~1 | study_ID,
      ~1 | subject_id,
      ~1 | biome_ID
    ),
    method = "REML",
    data = data_filtered,
    sparse = TRUE
  )

  return(model)
}

meta_models_disease <- list()

for (name in names(cleaned_datasets_pipelines)) {
  dat <- cleaned_datasets_pipelines[[name]]
  level <- ifelse(grepl("genus", name), "genus", "family")

  # Rho model
  model_rho <- run_meta_model_disease(dat, level = level, effect_size_col = "z", variance_col = "v")
  model_name_rho <- paste0("mod_rho_", name)
  meta_models_disease[[model_name_rho]] <- model_rho

  # Slope model
  model_slope <- run_meta_model_disease(dat, level = level, effect_size_col = "slope", variance_col = "ste")
  model_name_slope <- paste0("mod_slope_", name)
  meta_models_disease[[model_name_slope]] <- model_slope
}

meta_model_disease_summary <- extract_model_summary(meta_models_disease)

biome_base_colors <- c(
  "skin" = "#1b9e77",
  "gut" = "#d95f02",
  "oral" = "#7570b3",
  "respiratory" = "#e7298a",
  "vaginal" = "#66a61e"
)
library(colorspace)
label_colors <- c(
  "Skin: Healthy" = lighten(as.character(biome_base_colors["skin"]), amount = 0.3),
  "Gut: Healthy" = lighten(as.character(biome_base_colors["gut"]), amount = 0.3),
  "Gut: Disease" = as.character(biome_base_colors["gut"]),
  "Oral: Healthy" = lighten(as.character(biome_base_colors["oral"]), amount = 0.3),
  "Oral: Disease" = as.character(biome_base_colors["oral"]),
  "Respiratory: Healthy" = lighten(as.character(biome_base_colors["respiratory"]), amount = 0.3),
  "Respiratory: Disease" = as.character(biome_base_colors["respiratory"]),
  "Vaginal: Healthy" = lighten(as.character(biome_base_colors["vaginal"]), amount = 0.3),
  "Vaginal: Disease" = as.character(biome_base_colors["vaginal"])
)

orchard_plot(meta_models_disease$mod_slope_family.no_art, 
             mod = "biome_disease_label", 
             xlab = "Effect size Beta",
             group = "study_ID",  k = TRUE, g = FALSE, trunk.size = 1.5,)

orchard_plot(
  meta_models_disease$mod_slope_family.no_art,
  mod = "biome_disease_label",
  xlab = "Effect size Beta",
  group = "study_ID",
  k = TRUE,
  g = FALSE,
  trunk.size = 1.5
) +
  scale_color_manual(values = label_colors) +
  scale_fill_manual(values = label_colors)


plot_list_disease <- lapply(names(meta_models_disease)[c(3,4,7,8)], function(name) {
  model <- meta_models_disease[[name]]
  
  # Determine effect type and level
  effect <- if (str_detect(name, "slope")) "Beta" else "Z"
  level <- if (str_detect(name, "genus")) "Genus" else "Family"
  artifact <- if (str_detect(name, "no_art")) "No Artifacts" else "With Artifacts"
  
  # Create label
  plot_title <- paste(level, "-", effect, "-", artifact)
  
  # Make orchard plot
  p <- orchard_plot(
  model,
  mod = "biome_disease_label",
  xlab = paste("Effect size", effect),
  group = "study_ID",
  k = TRUE,
  g = FALSE,
  trunk.size = 1.5
) +
  scale_color_manual(values = label_colors) +
  scale_fill_manual(values = label_colors)+
  scale_x_discrete(labels = c("Skin", "Gut", "Gut (Dis.)", "Oral", "Oral (Dis.)", "Resp.", "Resp. (Dis.)", "Vag.", "Vag. (Dis.)"))
  
  return(p)
})

# Combine with patchwork
library(patchwork)
combined_plot <- wrap_plots(plotlist = plot_list_disease, ncol = 2)
combined_plot



```

